---
title: "02 Data Cleaning"
author: "Nathan Craig"
date: "6/25/2021"
output: 
    bookdown::html_document2: 
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
    code_download: true
---

# Introduction
The primary cleaning functions are incorporated into a custom function and moved to a separate R file. However, the mechanics of the transformations are displayed on this page.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE, warning=FALSE}
# Load necessary libraries
library(googlesheets4)
library(readr)
library(tidyverse)
library(janitor)
library(lubridate)
```

# Cleaning Script
First load the file and then clean the names. These operations were moved to a separate file. However, it is shown here to get a sense of what is involved.

```{r message=FALSE, warning=FALSE}

# Read in the sheet
df_inspect <- read_sheet("https://docs.google.com/spreadsheets/d/1im5VSi3bIEi13O8WQ56wEIXSyNEstbGMylXXgD9bAG0/edit#gid=1858227071",
                 sheet="Inspections",
                 col_names = TRUE,
                 col_types = "c") %>% 
  clean_names()
```

The following code is useful for isolating issues in the field names.
```{r message=FALSE, warning=FALSE, results='hide'}
df_inspect %>% 
  distinct(facility, .keep_all = TRUE) %>% 
  arrange(facility)
```

Here the following procedures were converted to a custom function and written to an R file. The procedures are illustrated here to show the transformations.
```{r message=FALSE, warning=FALSE}
df_inspect <- df_inspect %>% 
  # Removing a curious whitespace character that appears
  # in more recent facility names
  mutate(facility = str_replace_all(facility,
         pattern = " ",
         replacement = " ")) %>% 
  
  separate(.,
           col = facility,
           into = c("facility","inspection_date"),
           sep = "\\) - ") %>%
  separate(.,
           col = facility,
           into = c("facility","state"),
           sep = "\\(") %>% 

  # Deal with issues in the file names, state, and inspection date
  mutate(facility = str_trim(facility, side = "both"),
         state = str_trim(state, side = "both"),
         facility = str_replace_all(facility,
                            pattern = "^Adelanto ICE Processing Center-East",
                            replacement = "Adelanto ICE Processing Center - East"),
         facility = str_replace_all(facility,
                            pattern = "^Adelanto ICE Processing Center-West",
                            replacement = "Adelanto ICE Processing Center - West"),
         facility = str_replace_all(facility,
                            pattern = "^Allen Parish Detention Facility",
                            replacement = "Allen Parish Public Safety Complex"),
         facility = str_replace_all(facility,
                            pattern = "^Berks County Residential Center",
                            replacement = "Berks Family Residential Center"),
         inspection_date = replace(inspection_date, inspection_date=="Jan. 29 - 31, 2019", "Jan. 31, 2019"),
         facility = str_replace_all(facility,
                            pattern = "^Bristol County Jail$",
                            replacement = "Bristol County Jail and House of Correction"),
         facility = str_replace_all(facility,
                            pattern = "^Buffalo$",
                            replacement = "Buffalo Batavia Service Processing Center"),
         state = replace(state, state == "Batavia) Service Processing Center", "NY"),
         facility = str_replace_all(facility,
                            pattern = "^Calhoun County Jail",
                            replacement = "Calhoun County Correctional Center"),
         facility = str_replace_all(facility,
                            pattern = "^Clay County Justice Center",
                            replacement = "Clay County Jail"),
         facility = str_replace_all(facility,
                            pattern = "^Coastal Bend Detention Facility",
                            replacement = "Coastal Bend Detention Center"),
         state = replace(state, state == "David L. Moss Criminal Justice Center)", "OK"),
         facility = str_replace_all(facility,
                            pattern = "^Dodge County Detention Center",
                            replacement = "Dodge County Detention Facility"),
         facility = str_replace_all(facility,
                            pattern = "^Donald W. Wyatt Detention Center",
                            replacement = "Donald W. Wyatt Detention Facility"),
         facility = str_replace_all(facility,
                            pattern = "^Essex County Corrections Facility",
                            replacement = "Essex County Correctional Facility"),
         facility = str_replace_all(facility,
                            pattern = "^Farmville Detention Center$",
                            replacement = "Immigration Centers of America - Farmville"),
         facility = str_replace_all(facility,
                            pattern = "^Florence SPC",
                            replacement = "Florence Service Processing Center"),
         facility = str_replace_all(facility,
                            pattern = "^Houston CDF",
                            replacement = "Houston Contract Detention Facility"),
         state = replace(state, state == "Polk)", "TX"),
         facility = str_replace_all(facility,
                            pattern = "^Immigration Centers of America$",
                            replacement = "Immigration Centers of America - Farmville"),
         state = replace(state, state =="ICA", "VA"),
         inspection_date = replace(inspection_date,
                                   inspection_date == "Farmville Detention Center (FDC) (VA",
                                   "Feb. 24, 2021"),
         # This one picks a value in the facility col and changes a value in the state col
         state = replace(state, facility == "Immigration Centers of America - Farmville",
                         "VA"),
         facility = str_replace_all(facility,
                            pattern = "^Joe Corley Detention Facility",
                            replacement = "Joe Corley Processing Center"),
         facility = str_replace_all(facility,
                            pattern = "^Karnes County Residential Center",
                            replacement = "Karnes County Family Residential Center"),
         facility = str_replace_all(facility,
                            pattern = "^Krome SPC",
                            replacement = "Krome Service Procesing Center"),
         facility = str_replace_all(facility,
                            pattern = "^Krome Special Processing Center",
                            replacement = "Krome Service Procesing Center"),
         state = replace(state, state =="SPC)", "FL"),
         facility = str_replace_all(facility,
                            pattern = "^Mesa Verde Detention Facility",
                            replacement = "Mesa Verde ICE Processing Facility"),
         facility = str_replace_all(facility,
                            pattern = "^Northwest Contract Detention Center",
                            replacement = "Northwest ICE Processing Center"),
         facility = str_replace_all(facility,
                            pattern = "^Northwest Detention Center",
                            replacement = "Northwest ICE Processing Center"),
         facility = str_replace_all(facility,
                            pattern = "^Okmulgee County Jail-Moore Detention Facility",
                            replacement = "Okmulgee County Jail - Moore Detention Facility"),
         facility = str_replace_all(facility,
                            pattern = "^Orange County Jail",
                            replacement = "Orange County Correctional Facility"),
         facility = str_replace_all(facility,
                            pattern = "^Otay Mesa Detention Facility",
                            replacement = "Otay Mesa Detention Center"),
         facility = str_replace_all(facility,
                            pattern = "^Prarieland Detention Center",
                            replacement = "Prairieland Detention Center"),
         facility = str_replace_all(facility,
                            pattern = "^Richwood Correcrtional Center",
                            replacement = "Richwood Correctional Center"),
         facility = str_replace_all(facility,
                            pattern = "^Rio Grande Processing Center",
                            replacement = "Rio Grande Detention Center"),
         facility = str_replace_all(facility,
                            pattern = "^Robert A. Deyton Correctional Center",
                            replacement = "Robert A. Deyton Detention Facility"),
         facility = str_replace_all(facility,
                            pattern = "^Robert A. Deyton Detention Center",
                            replacement = "Robert A. Deyton Detention Facility"),
         facility = str_replace_all(facility,
                            pattern = "^South Texas Detention Complex",
                            replacement = "South Texas ICE Processing Center"),
         facility = str_replace_all(facility,
                            pattern = "^South Texas Processing Center",
                            replacement = "South Texas ICE Processing Center"),
         facility = str_replace_all(facility,
                            pattern = "^Strafford County Corrections",
                            replacement = "Strafford County Department of Corrections"),
         inspection_date = replace(inspection_date, state =="CO)- Mar. 31, 2021", "Mar. 31, 2021"),
         state = replace(state, state == "CO)- Mar. 31, 2021", "CO"),
         facility = str_replace_all(facility,
                            pattern = "^T. Don Hutto Detention Center$",
                            replacement = "T. Don Hutto Residential Center"),
         facility = str_replace_all(facility,
                            pattern = "Tulsa County Jail - David L. Moss Criminal Jutice Center",
                            replacement = "David L. Moss Criminal Justice Center"),
         facility = str_replace_all(facility,
                            pattern = "^Tulsa County Jail$",
                            replacement = "David L. Moss Criminal Justice Center"),
         state = replace(state, state == "David L. Moss Justice Center)", "OK"),
         facility = str_replace_all(facility,
                            pattern = "^Washoe County Detention Centerr",
                            replacement = "Washoe County Detention Center"),
         facility = str_replace_all(facility,
                            pattern = "^Webb County Detention Facility",
                            replacement = "Webb County Detention Center"),
         )
```



