# Solitary

```{r setup-6, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(scipen=999)
```

```{r load-libraries-6, message=FALSE}
# Plotting
library(ggplot2)
library(gghighlight)
library(RColorBrewer)

# Tables
library(kableExtra)

load("data/data.RData")
```

```{r}
df_solitary_sum <- df_solitary %>% 
  group_by(facility) %>% 
  filter(segregation_type == "total_segregation") %>% 
  summarise(total_segregation_facility = sum(segregation_count, na.rm = TRUE)) %>% 
  arrange(desc(total_segregation_facility)) %>% 
  ungroup

df_solitary_sum <- left_join(df_solitary_sum, df_facility_n_inspect) %>% 
  mutate(total_segregation_by_inspection = round(total_segregation_facility / n_inspections, digits = 0)) %>% 
  relocate(fac_operator:n_inspections, .after= facility)

m_segregation_fac_by_insp <- mean(df_solitary_sum$total_segregation_by_inspection)
sd_segregation_fac_by_insp <- sd(df_solitary_sum$total_segregation_by_inspection)
outlier_segregation_fac_by_insp <- m_segregation_fac_by_insp + 2*sd_segregation_fac_by_insp
```

ICE PBNDS defines a "Special Management Unit" as "housing unit for detainees in administrative or disciplinary segregation" and defines segregation as "[c]onfinement in an individual cell isolated from the general population; for administrative, disciplinary, or protective reasons" [@ice2016, 473, 474]. These are anodyne euphemisms for solitary confinement. The G-324A SIS form records information on four kinds of solitary confinement:

-   administrative
-   disciplinary
-   medical
-   mental health

During the `r inspect_sample_interval` months covering the period between `r format(inspect_earliest, format= "%B %e, %Y")` and `r format(inspect_latest, format= "%B %e, %Y")`, there were `r segregation_sum` instances of solitary reported (Tables \@ref(tab:solitary-summary) and \@ref(tab:solitary-by-facility)).

There was an average of `r round(m_segregation_fac_by_insp, digits = 2)` (SD=`r round(sd_segregation_fac_by_insp, digits =2)`) instances of segregation per facility by inspection. Six facilities have significantly higher than average number of solitary instances reported per inspection. Of these facilities two are managed by ICE, one is managed by the Henderson Police Department, two are operated by CoreCivic, and one is managed by LaSalle Corrections.

Roughly half of the total instances of solitary were administrative segregation (`r segregation_admin_perc`%, n = `r segregation_admin_sum`) while disciplinary (`r segregation_disciplinary_perc`%, n = `r segregation_disciplinary_sum`) and medical (`r segregation_medical_perc`%, n =`r segregation_medical_sum`) segregation made up nearly the other half. Mental health segregation (`r segregation_mental_health_perc`%, n = `r segregation_mental_health_sum`) comprised a relatively small proportion of the total number of solitary instances and still surpassed 1000 instances.

```{r solitary-summary}
df_solitary %>% 
  group_by(segregation_type) %>% 
  summarise(`Total Solitary by Type` = sum(segregation_count, na.rm = TRUE)) %>% 
  ungroup() %>% 
  kable(caption = "Total Solitary by Type",
        col.names = c("Solitary Type", "Total Solitary Type")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))


# df_solitary %>% 
#   group_by(facility) %>% 
#   summarise(total_segregation = sum(segregation_count, na.rm = TRUE)) %>% 
#   arrange(desc(total_segregation)) %>% 
#   ungroup() %>% 
#   kable(caption = "Total Solitary by Facility",
#         col.names = c("Facility", "Total Solitary by Facility")) %>% 
#   kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
#   scroll_box(height = "300px")
```

```{r solitary-by-facility}
color.me <- which(df_solitary_sum$total_segregation_by_inspection > outlier_segregation_fac_by_insp)

df_solitary_sum %>% 
  group_by(facility) %>% 
  arrange(desc(total_segregation_facility)) %>%
  select(-n_inspections) %>% 
  ungroup() %>% 
  kable(caption = "Total solitary by facility, with facilities having significnatly greater than average number of solitary instances per inspection shown in bold.",
        col.names = c("Facility", "Facility Operator",  "Total Solitary by Facility", "Total Solitary by Inspection")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  row_spec(color.me, bold = TRUE) %>% 
  scroll_box(height = "400px")
```

```{r histogram-solitary-by-inspection, fig.cap="Histogram of solitary instances reported by inspection highlighting facilities that have significantly higher than average number of solitary instances reported per inspection."}
df_solitary_sum %>% 
  ggplot(aes(x = total_segregation_by_inspection, fill = facility))+
  geom_histogram()+
  gghighlight(max(total_segregation_by_inspection) > outlier_segregation_fac_by_insp, use_direct_label = FALSE)+
  labs(title = "Histogram solitary instances reported by inspection",
       subtitle = "Facilities with significantly higher than average number of solitary instances reported\nper inspection highlighted.",
       fill = "Facility")+
  xlab("Solitary by Inspection")+
  ylab("Number of Facilities")
```

```{r solitary-greater-than-average-facility, fig.cap="Facilities in the sample with greater than average number of solitary instances reported per inspection."}
df_solitary_sum %>% 
  filter(total_segregation_by_inspection > outlier_segregation_fac_by_insp) %>% 
  ggplot(aes(x = total_segregation_by_inspection, y=reorder(facility, -total_segregation_by_inspection, sum)))+
  geom_bar(stat="identity")+

  xlab("Solitary by Inspection")+
  ylab("Facility")
```

## Solitary Over 60 Days

```{r}
df_seg60_sum <- df_324 %>% 
  select(facility, 
         fac_operator, 
         admin_seg_60_ice,
         disc_seg_60_ice) %>% 
  mutate(total_seg_60 = admin_seg_60_ice+disc_seg_60_ice)
  

df_seg60_sum <- left_join(df_seg60_sum, df_facility_n_inspect) %>%
  drop_na %>% 
  # filter(seg60_type == "total_seg_60") %>% 
  mutate(total_seg60_by_inspection = round(total_seg_60 / n_inspections, digits = 0)) %>% 
  relocate(fac_operator:n_inspections, .after= facility) %>% 
  arrange(desc(total_seg60_by_inspection))

# Variables
sum_seg60 <- sum(df_seg60_sum$total_seg_60)
m_seg60_insp <- round( mean(df_seg60_sum$total_seg60_by_inspection), digits = 2)
sd_seg60_insp <- round( sd(df_seg60_sum$total_seg60_by_inspection), digits = 2)

admin_seg60_sum <- sum(df_324$admin_seg_60_ice, na.rm = TRUE)
admin_seg60_perc <- round((sum(df_324$admin_seg_60_ice, na.rm = TRUE)/sum_seg60)*100, digits = 2)

disc_seg60_sum <- sum(df_324$disc_seg_60_ice, na.rm = TRUE)
disc_seg60_perc <- round((sum(df_324$disc_seg_60_ice, na.rm = TRUE)/sum_seg60)*100, digits = 2)
```

```{r}
df_seg60_sum_facility <- df_seg60_sum %>% 
  group_by(facility, fac_operator) %>% 
  summarise(sum_seg60_insp = sum(total_seg60_by_inspection)) %>% 
  arrange(desc(sum_seg60_insp))

m_seg60_insp <- mean(df_seg60_sum_facility$sum_seg60_insp)
sd_seg60_insp <- sd(df_seg60_sum_facility$sum_seg60_insp)
outlier_seg60_insp <- m_seg60_insp + 2*sd_seg60_insp
```

The G-324A SIS form includes line items for annual number of administrative and disciplinary segregation stays over 60 days. ICE PBNDS Appendix 3.1.A, Offense Categories I.B.3 defines up to 60 days as the maximum length of disciplinary segregation [@ice2016, 224]. It is unclear why form G-324A has an entry for reporting uses of disciplinary solitary exceeding the 60 day maximum defined by PBNDS.

Based on the inspections reviewed there were `r sum_seg60` instances of segregation \> 60 days reported. Of these cases, `r admin_seg60_perc`% (n=`r admin_seg60_sum`) involved administrative solitary and `r disc_seg60_perc`% (n=`r disc_seg60_sum`) involved disciplinary solitary.

Within this sample of inspections, there were on average `r round(m_seg60_insp, digits = 2)` (SD=`r round(sd_seg60_insp, digits = 2)`) instances of segregation greater than 60 days per inspection. There are five facilities that have significantly higher than average number of incidents of solitary greater than 60 days.

Of these, one is operated by MTC, three by the GEO Group, and one by CoreCivic. MTC's Imperial Regional Detention Facility stands out with more than double the next two highest ranking facilities regarding solitary \> 60 days.

```{r}
color.me <- which(df_seg60_sum_facility$sum_seg60_insp > outlier_seg60_insp)

df_seg60_sum_facility %>% 
 
  kable(caption = "Total solitary >60 days by facility, with facilities having significnatly greater than average number of solitary >60 days instances per inspection shown in bold.",
        col.names = c("Facility", "Facility Operator",  "Total Solitary > 60 days by Facility by Inspection")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  row_spec(color.me, bold = TRUE) %>% 
  scroll_box(height = "400px")
```

```{r solitary-60-facet}

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(5, "Paired")
plot_colors[5] <- "#000000"

# Call the dataframe and select cols
df_324 %>% 
  select(id,
         facility,
         state,
         date,
         fac_operator,
         admin_seg_60_ice,
         disc_seg_60_ice) %>% 
  drop_na() %>% 
  
  # Need the rowwise function to compute a row-at-a-time
  # in the following mutate function
  rowwise(id) %>% 
  
  # Generate total col
  mutate(total_seg_60 = sum(c_across(
       admin_seg_60_ice:
       disc_seg_60_ice
  ), na.rm = TRUE)) %>%
  
  # Make tidy and filter
  pivot_longer(cols = admin_seg_60_ice:disc_seg_60_ice,
               names_to = "segregation_60_type",
               values_to = "segregation_60_count") %>%
  filter(segregation_60_type %in% c("admin_seg_60_ice", "disc_seg_60_ice")&
           segregation_60_count > 0) %>%
  
  # Initiate the plot and sort by sum
  ggplot(aes(x = segregation_60_count,
             y=reorder(fac_operator, segregation_60_count, sum),
             fill=segregation_60_type))+
  geom_bar(stat = "identity")+
  
  # Set the color fill
  scale_fill_brewer(type = "qual",
                    palette = "Paired",
                    name = "Segregation > 60 Type",
                    labels = c("Administrative",
                               "Disciplinary"))+
  labs(title= "Segregation > 60 Days by Facility Operator",
        x = "Segregation > 60 Days Count",
        y = "Facility Operator")+
  theme(legend.position = "bottom")
```
