# Assaults

```{r setup-4, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
options(scipen=999)
```

```{r load-libraries-4, message=FALSE}
library(tidyverse)
# Plotting
library(ggplot2)
library(gghighlight)
library(RColorBrewer)

# Tables
library(kableExtra)

load("data/data.RData")
```

The ICE G-324A Significant Incident Summary sheet contains information on four categories of assaults:

-   Detained Person on Staff with Serious Injury
-   Detained Person on Staff without Serious Injury
-   Detained Person on Detained Person with Serious Injury
-   Detained Person on Detained Person without Serious Injury

Out of the `r inspect_proc_n` inspections that occurred at `r facility_n` facilities between `r format(inspect_earliest, format= "%B %e, %Y")` and `r format(inspect_latest, format= "%B %e, %Y")`, there were `r assault_sum` assaults (Table \@ref(tab:assault-summary-by-type)).

Over the course of `r inspect_sample_interval` months represented by the inspections in this sample, there were on average `r round(assault_sum/inspect_sample_interval, digits=0)` assaults per month.The majority of those assaults occurred between detained persons and did not result in serious injury (86%, n=2926). The SIS form does distinguish if an assault on a person detained by ICE was by another person detained by ICE or by a person detained under a corrections context. This is an important detail because many facilities are mixed use wherein individuals detained for civil immigration purposes are in corrections facilities.

```{r assault-summary-by-type}
df_assaults %>% 
  group_by(assault_type) %>% 
  summarise(`Total Assaults by Type` = sum(assault_count, na.rm = TRUE)) %>% 
  ungroup() %>% 
  kable(caption = "Total Assaults by Type",
        col.names = c("Assault Type", "Total Assault Type")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))
```

```{r df-assaults-sum}
df_assaults_sum <- df_assaults %>% 
  group_by(facility) %>% 
  filter(assault_type == "total_assaults") %>% 
  summarise(total_assaults = sum(assault_count, na.rm = TRUE)) %>% 
  arrange(desc(total_assaults)) %>% 
  ungroup()
  
df_assaults_sum <- left_join(df_assaults_sum, df_facility_n_inspect) %>% 
  mutate(total_assaults_by_inspection = round(total_assaults / n_inspections, digits = 0))


assaults_fac_by_insp_m <- mean(df_assaults_sum$total_assaults_by_inspection)
assaults_fac_by_insp_sd <- sd(df_assaults_sum$total_assaults_by_inspection)
assault_fac_by_insp_outlier <- assaults_fac_by_insp_m+(2*assaults_fac_by_insp_sd)
```

The mean number of assaults by facility per inspection is `r round(assaults_fac_by_insp_m, digits=0)` (SD=`r round(assaults_fac_by_insp_sd, digits=2)`) (Figure \@ref(fig:assaults-bar-by-facility) and Table \@ref(tab:assault-summary-by-facility)). Based on this mean, five facilities have significantly greater than average number of assaults per inspection (\> `r round(assault_fac_by_insp_outlier, digits=2)` assaults). Among the facilities with significantly higher than average number of assaults per inspection three are managed by The Geo Group; one is managed by LaSalle Corrections; and one is a Service Processing Center managed by ICE.

```{r assaults-bar-by-facility, fig.cap="Bar plot of facilities with greater than average number of assaults reported per inspection.", fig.height=6}
threshold <- assaults_fac_by_insp_m + assaults_fac_by_insp_sd *2

df_assaults_sum %>% 
  group_by(facility, fac_operator) %>% 
  filter(total_assaults_by_inspection > assaults_fac_by_insp_m) %>% 
  ggplot(aes(x = total_assaults_by_inspection, y = reorder(facility, total_assaults_by_inspection, sum)))+
  geom_bar(stat = "identity")+
  gghighlight(max(total_assaults_by_inspection) > threshold)+
  labs(title = "Facilities with > μ Assaults per Inspection.",
       subtitle = "Facilities with significantly higher than average number\nof assaults are dark grey.")+
  xlab("Assaults Per Inspection")+
  ylab("Facility")
```

```{r assault-summary-by-facility}
color.me <- which(df_assaults_sum$total_assaults_by_inspection > assault_fac_by_insp_outlier)

df_assaults_sum %>% 
  arrange(desc(total_assaults_by_inspection)) %>% 
  select(-n_inspections) %>%
  relocate(fac_operator, .after = facility) %>% 
  kable(caption = "Assaults by facility, facilities with significantly greater than average number of assaults by inspection shown in bold.",
        col.names = c("Facility", "Facility Operator", "Total Assaults", "Assaults by Inspections")) %>%
  kable_styling(c("hover", "striped", "condensed", "responsive"), 
                font_size = 12) %>% 
  row_spec(color.me, bold = TRUE) %>%
  scroll_box(height = "400px")
```

## Discussion

Conspicuously absent from the four assault types listed on ICE's G-324A SIS form are assaults on detained persons by staff. This omission is deeply concerning because prior news stories and review of the inspection forms themselves show that staff assaults on detained persons do occur. However, these are generally described as "use of force" incidents and are not tallied as assaults. Use of force incidents are described in Section \@ref(use-of-force).
