# Cause of Death

```{r setup-10, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

options(scipen=999)
```

```{r load-libraries-10}
library(stringr)
library(janitor)

# Plotting
library(ggplot2)
library(RColorBrewer)

# Tables
library(kableExtra)

load("data/data.RData")
# # Read data
# source("function_load_324.R", local = knitr::knit_global())
```

Form G-324A contains fields for reporting cause of death (COD) for persons incarcerated by ICE and persons incarcerated not by ICE. The non-ICE COD reporting is presumably for corrections facilities where ICE has civil immigration detention bed contracts.

```{r}
df_cod %>% 
  filter(!str_detect(cod_type, "not_ice")) %>% 
  group_by(cod_type) %>% 
  summarise(total = sum(cod_count, na.rm = TRUE)) %>% 
  arrange(desc(total)) %>% 
  ungroup() %>% 
  adorn_totals("row") %>% 
  kable(caption = "COD Totals by Category") %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))
```

```{r}
df_cod %>% 
  filter(!str_detect(cod_type, "not_ice"),
         cod_count>0,
         cod_type == "total_cod") %>% 
  left_join(., df_facility_n_inspect) %>% 
  mutate(cod_count_insp = round(cod_count/n_inspections, digits = 0)) %>% 
  select(facility, fac_operator, cod_count, cod_count_insp) %>% 
  group_by(facility) %>% ungroup() %>% 
  arrange(desc(cod_count)) %>% 
  kable(caption = "COD by Facility",
        col.names = c("Facility", "Facility Operator", "COD Total", "COD Total by Inspection")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"),
                font_size = 12) %>% 
  scroll_box(height = "400px")
```

## Suicide as Cause of Death

Bar plot of ICE and non-ICE deaths by suicide plotted by facility as of `r Sys.Date()`.

```{r suicide-cod-plot}
 
# df_324 %>%
#   select(id,
#          facility,
#          state,
#          date,
#          fac_operator,
#          cod_suicide_ice,
#          cod_suicide_not_ice) %>%
#   drop_na() %>%
#   filter(cod_suicide_ice>0|cod_suicide_not_ice >0) %>% 
#   mutate(cod_suicide_total = cod_suicide_ice + cod_suicide_not_ice) %>% 
  
  
  
df_cod %>% 
  left_join(., df_facility_n_inspect) %>%
  filter(!str_detect(cod_type, "not_ice"),
        cod_count >0,
        cod_type == "total_cod") %>% 
  mutate(cod_count_insp = round(cod_count/n_inspections, digits = 0)) %>% 
  select(facility, fac_operator, cod_count, cod_count_insp) %>% 
  
  
  # Here one could call datatable() or kable()
  # to get a list of the facilities and count
  
  # Initiate the plot
  ggplot(aes(x=cod_count_insp, y=reorder(facility, cod_count, sum)))+
  
  
  geom_bar(stat = "identity")+
  # scale_x_continuous(breaks = c(1,2))+
  labs(title = "Cause of Death Suicide",
       subtitle = "Includes ICE and non-ICE Deaths by Suicide")+
  xlab("COD Suicide Total")+
  ylab("Facility")
```
