# Disciplinary

```{r setup-5, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(scipen=999)
```

```{r load-libraries-5, message=FALSE}
# Plotting
library(ggplot2)
library(RColorBrewer)

# Tables
library(kableExtra)

load("data/data.RData")
```

```{r}
df_discipline_wide <- df_324_inc %>% 
  select(id, 
         facility, 
         inspection_date, 
         date, 
         disciplinary_infractions, 
         disciplinary_infractions_guilty, 
         disciplinary_appeals, 
         disciplinary_appeals_found_in_favor_of_detainee,
         sanctions_over_60_days) %>% 
  left_join(., df_facility_n_inspect) %>% 
  relocate(fac_operator, .after = facility) %>%
  relocate(n_inspections, .after = date)
```

```{r}
df_discipline_sum <- df_discipline %>%
  group_by(facility) %>%
 filter(disciplinary_type == "disciplinary_infractions") %>%
 summarise(total_disciplinary_facility = sum(disciplinary_count, na.rm = TRUE)) %>%
 arrange(desc(total_disciplinary_facility)) %>%
 ungroup

df_discipline_sum <- left_join(df_discipline_sum, df_facility_n_inspect) %>%
  mutate(total_disciplinary_by_inspection = round(total_disciplinary_facility / n_inspections, digits = 0)) %>%
 relocate(fac_operator:n_inspections, .after= facility)


discipline_fac_by_insp_m <- mean(df_discipline_sum$total_disciplinary_by_inspection)
discipline_fac_by_insp_sd <- sd(df_discipline_sum$total_disciplinary_by_inspection)
discipline_fac_by_insp_outlier <- discipline_fac_by_insp_m + 2*discipline_fac_by_insp_sd
```

As a matter of law, ICE detention is not punitive it is purely administrative civil holding to ensure that individuals are present for their immigration proceedings [@ice2016, i; @schriro2009, 2-3]. Despite that fact, punitive disciplinary actions are logged on ICE's G-324A SIS form. The form also lists the number of disciplinary fractions found guilty, appeals, and appeals in favor of detained persons.

During the `r inspect_sample_interval` months covering the period between `r format(inspect_earliest, format= "%B %e, %Y")` and `r format(inspect_latest, format= "%B %e, %Y")`, there were `r disciplinary_sum` disciplinary infractions issued by facility operators (Table \@ref(tab:total-disciplinary-type)). There were an average of `r round(discipline_fac_by_insp_m, digits = 2)` (SD = `r round(discipline_fac_by_insp_sd, digits = 2)`) disciplinary infractions per facility by inspection (Figure \@ref(fig:disciplinary-bar-facility) and Table \@ref(tab:disciplinary-by-facility)).

Four facilities have significantly greater than average number of disciplinary infractions per inspection. Of the facilities with significantly greater than average number of disciplinary infractions issued per inspection two are managed by CoreCivic, one is managed by Immigration Centers of America, and one is a service processing center managed by ICE.

```{r total-disciplinary-type}
df_discipline %>% 
  group_by(disciplinary_type) %>% 
  summarise(`Total Disciplinary by Type` = sum(disciplinary_count, na.rm = TRUE)) %>% 
  ungroup() %>% 
  kable(caption = "Total Disciplinary by Type",
        col.names = c("Discipline Type", "Total Discipline Type")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))
```

```{r disciplinary-bar-facility, fig.cap="Bar plot of facilities with greater than average number of Disciplinary infractions filed by inspection.",fig.height=6}
threshold <- discipline_fac_by_insp_m + discipline_fac_by_insp_sd *2

df_discipline_sum %>% 
  group_by(facility, fac_operator) %>% 
  filter(total_disciplinary_by_inspection > discipline_fac_by_insp_m) %>% 
  ggplot(aes(x = total_disciplinary_by_inspection, y = reorder(facility, total_disciplinary_by_inspection, sum)))+
  geom_bar(stat = "identity")+
  gghighlight(max(total_disciplinary_by_inspection) > threshold)+
  labs(title = "Facilities with > μ Disciplinary per Inspection Reported.")+
  xlab("Disciplinary Per Inspection")+
  ylab("Facility")
```

```{r disciplinary-by-facility}
color.me <- which(df_discipline_sum$total_disciplinary_by_inspection > discipline_fac_by_insp_outlier)

df_discipline_sum %>% 
  group_by(facility) %>% 
  # summarise(total_disciplinary = sum(disciplinary_count, na.rm = TRUE)) %>% 
  arrange(desc(total_disciplinary_facility)) %>%
  select(-n_inspections) %>% 
  ungroup() %>% 
  kable(caption = "Total discipline by facility, with facilities having significnatly greater than average number of disciplinary infractions issued per inspection shown in bold.",
        col.names = c("Facility", "Facility Operator",  "Total Discipline by Facility", "Total Disciplinary by Inspection")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"),
                font_size = 12) %>% 
  row_spec(color.me, bold = TRUE) %>% 
  scroll_box(height = "400px")
  
```
