# Medical

```{r setup-9, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
options(scipen=999)
```

```{r load-libraries-9, message=FALSE}
# Plotting
library(ggplot2)
library(RColorBrewer)

# Tables
library(kableExtra)

load("data/data.RData")
```

## Medical Observation

```{r medical-observation-summary}
df_observation %>% 
  group_by(medical_mental_observation_type) %>% 
  summarise(`Total Observation by Type` = sum(medical_mental_observation_count, na.rm = TRUE)) %>% 
  ungroup() %>% 
  kable(caption = "Total Medical/Mental Observation by Type",
        col.names = c("Type", "Total")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))

```

```{r medical-observation-fac-insp}
df_observation %>% 
  filter(medical_mental_observation_type == "total_mental_medical_observation") %>%
  group_by(facility) %>% 
  summarise(total_observation = sum(medical_mental_observation_count, na.rm = TRUE)) %>% 
  
  left_join(., df_facility_n_inspect) %>% 
  mutate(total_observation_insp = round(total_observation/n_inspections, digits = 0)) %>% 
  arrange(desc(total_observation)) %>% 
  select(facility, fac_operator, total_observation, total_observation_insp) %>% 
  ungroup() %>% 
  kable(caption = "Total Medical/Mental Observations by Facility",
        col.names = c("Facility", "Facility Operator", "Total Observation", "Total Observation by Inspection")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "400px")
```

## Infections Disease

```{r infectious-disease-summary}

df_infectious %>% 
  group_by(infectious_disease_type) %>% 
  summarise(`Total by Type` = sum(infectious_disease_count, na.rm = TRUE)) %>% 
  ungroup() %>% 
  kable(caption = "Total by Type",
        col.names = c("Type", "Total")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))
```

```{r infectious-disease-by-fac-inspect}
df_infectious %>% 
  filter(infectious_disease_type == "infections_disease_confirmed") %>% 
  group_by(facility) %>% 
  summarise(total_infectious_confirmed = sum(infectious_disease_count, na.rm = TRUE)) %>%
  left_join(., df_facility_n_inspect) %>% 
  mutate(infectious_disease_confirmed_insp = round(total_infectious_confirmed/n_inspections, digits = 0)) %>% 
  select(facility, fac_operator, total_infectious_confirmed, infectious_disease_confirmed_insp) %>%
  arrange(desc(total_infectious_confirmed)) %>% 
  ungroup() %>% 
  kable(caption = "Total Confirmed by Facility",
        col.names = c("Facility", "Facility Operator", "Total Confirmed", "Total Confirmed by Inspection")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "300px")
```

## Medical and Mental Health Referrals

```{r referrals-summary}
df_referrals %>% 
  group_by(referral_type) %>% 
  summarise(`Total Referral by Type` = sum(referral_count, na.rm = TRUE)) %>% 
  ungroup() %>% 
  kable(caption = "Total Referrals by Type",
        col.names = c("Type", "Total")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))
```

```{r referrals-facility-insp}
df_referrals %>% 
  filter(referral_type == "total_referrals") %>% 
  group_by(facility) %>% 
  summarise(total_referral = sum(referral_count, na.rm = TRUE)) %>% 
  left_join(., df_facility_n_inspect) %>% 
  mutate(total_referral_insp = round(total_referral/n_inspections, digits = 0)) %>% 
  select(facility, fac_operator, total_referral, total_referral_insp) %>% 
  arrange(desc(total_referral)) %>% 
  ungroup() %>% 
  kable(caption = "Total Referrals by Facility",
        col.names = c("Facility", "Facility Operator", "Total Referrals", "Total Referrals by Inspection")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "300px")
```

## Sick Call Requests and Encounters

```{r sick-calls-summary}
df_sick_call %>% 
  group_by(sick_call_type) %>% 
  summarise(`Total Sick Call by Type` = sum(sick_call_count, na.rm = TRUE)) %>% 
  ungroup() %>% 
  kable(caption = "Total Sick Call by Type",
        col.names = c("Type", "Total")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))
```

```{r sick-calls-fac-insp}
df_sick_call %>% 
  filter(sick_call_type == "total_sick_call") %>% 
  group_by(facility) %>% 
  summarise(total_sick_call = sum(sick_call_count, na.rm = TRUE)) %>% 
  arrange(desc(total_sick_call)) %>% 
  left_join(., df_facility_n_inspect) %>% 
  mutate(total_sick_call_insp = round(total_sick_call/n_inspections, digits = 0)) %>% 
  select(facility, fac_operator, total_sick_call, total_sick_call_insp) %>% 
  ungroup() %>% 
  kable(caption = "Total Sick Call Requests/Encounters by Facility",
        col.names = c("Facility", "Facility Operator", "Total Sick Call", "Total Sick Call by Inspection")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "300px")
```

## Suicide Attempts and Watches

The following shows Suicide Attempts and Suicide Watches based on the status of the data as of `r Sys.Date()`. The numbers are subject to change.

```{r suicide-summary}
df_suicide %>% 
  group_by(suicide_type) %>% 
  summarise(`Total Suicide Attempt or Watch by Type` = sum(suicide_count, na.rm = TRUE)) %>% 
  ungroup() %>% 
  kable(caption = "Total Suicide Attempt or Watch by Type",
        col.names = c("Type", "Total")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))
```

```{r}
df_suicide %>% 
  filter(suicide_type == "total") %>% 
  group_by(facility) %>% 
  summarise(total_incidents = sum(suicide_count, na.rm = TRUE)) %>% 
  left_join(., df_facility_n_inspect) %>% 
  mutate(total_incidents_insp = round(total_incidents/n_inspections, digits = 0)) %>% 
  select(facility, fac_operator, total_incidents, total_incidents_insp) %>% 
  arrange(desc(total_incidents)) %>% 
  ungroup() %>% 
  kable(caption = "Total Suicide Attempt or Watch by Facility",
        col.names = c("Facility", "Facility Operator", "Total Incidents", "Total Incidents by Inspection")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "400px")
```

## Hunger Strikes

There were `r hunger_strike_sum` at `r hunger_strike_facility_n` facilities, Table \@ref(tab:hunger-strike-summary) provides a breakdown of hunger strikes by facility.

```{r hunger-strike-summary}
df_hunger_strike %>%
  group_by(facility) %>% 
  summarise(total_hunger_strike = sum(hunger_strikes, na.rm = TRUE)) %>% 
  left_join(., df_facility_n_inspect) %>% 
  select(facility, fac_operator, total_hunger_strike) %>% 
  
  arrange(desc(total_hunger_strike)) %>% 
  ungroup() %>% 
  kable(caption = "Total Hunger Strikes by Facility",
        col.names = c("Facility", "Facility Operator", "Total Hunger Strikes")) %>%
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "300px")
```
