# (APPENDIX) Appendix {.unnumbered}

# Reading in SIS Forms

```{r setup-appendix, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
options(scipen=999)
```

```{r load-libraries-appendix, message=FALSE, warning=FALSE}
# Plotting
library(ggplot2)
library(RColorBrewer)

# Tables
library(kableExtra)

load("data/data.RData")
```

The code for this report is specific to data logged in the G-324A-19 form and the incident summaries portion of that form. The incident summary portion of this form is probably the most complex data structure in the project and also contains the largest volume of quantitative information. Therefore, it seemed like a good place to start with an initial proof of concept.

At the outset, there were issues reading in the Google Sheet. Several of the numeric columns read into R as lists which was undesirable. In the Google Sheet, Craig set the columns to `plain text` rather than `auto` and the data are read into R as character columns. This seemed to resolve the issue for a time, and it was possible to use the `type_convert()` function from the `readr` library [@R-readr] to convert these columns to numbers. The `col_types` argument of the `read_sheet` function of the `googlesheets4` library [@R-googlesheets4] was used to explicitly set the column types to character. After applying the `clean_facility_names` custom function, additional transformations like converting facility to a factor are applied.

There were some issues using `lubridate` [@R-lubridate] to wrangle the month and year fields to a proper date. Craig was not able to discern the cause of the issue. However, changing the month format on the incident sheet from abbreviated to fully written out months solved the issue. Those changes were made in the master Google Sheet and the data were called using the `googlesheets4` library.

At present `r inspect_n` SIS and inspection cover letter combinations out of approximately 300 inspections are complete. The completed inspections range in time from `r inspect_earliest` to `r inspect_latest`. Students are now working on older SIS forms which are structured differently. The preliminary EDA in this document is restricted to the more current SIS forms which were first used in May 2019.

## Summary Tables

Summary tables help to provide an overview of how many instances occur within particular category of data or how frequent a particular issue is recorded at a given facility. Summary tables are generated using a combination of `group_by` with `summarize` with the results piped to `kable` and `kableExtra`. This produces results similar to a "pivot table" from Excel.

## Facet Plots

Facet plots partition each portion of a plot into a matrix of panels. These plots are produced for several categories of data where each panel represents a facility for a given variable reported in the SIS form. Within each category, columns are pivoted longer and plotted by date. Graphing is done using the `ggplot2` library [@R-ggplot2] with the `facet_wrap` function to provide a means to compare multiple facilities simultaneously. Such a plot can help identify trends and guide more specific questions.

# Data Cleaning

Anyone familiar working with ICE data knows that the agency is notorious for releasing dirty data filled with pesky errors. For example, consider the Krome Service Processing Center, the very first immigration processing center established in the 1980's after the closure of Ellis Island in 1954. Of all the facilities the agency should name consistently, Krome the very first detention center of the contemporary system, one would expect ICE to get right. However, the facility inspections page lists the following names for the Krome Service Processing Center:

-   Krome Special Processing Center (SPC) (FL) - Feb. 10, 2021
-   Krome Service Procesing Center (FL) - Feb. 13, 2020
-   Krome SPC (FL) - Feb. 14, 2019

Every instance is different. In one case "Special" is replaced for "Service." In another case the word processing is misspelled "Procesing," while in a third, the title is abbreviated. This is one example of many. In other cases facilities are sometimes named "Detention Center" rather than "Detention Facility," etc. Assembling the data into a useful table that can be sorted, filtered, and summarized requires standardizing the names.

From early on, Craig decided to assemble the data as reported, errors and all, and perform standardization programmatically. Doing this facilitates reproducibility. The project's primary cleaning functions are incorporated into a custom function `clean_facility_names()` and moved to a [separate R file](https://github.com/n8craig/inspections/blob/main/function_clean_facility_names.R). This allows more than 100 lines of code to be collapsed into a single function call. However, for display purposes, the lines of code that make up the transformations are detailed on this page.

```{r message=FALSE, warning=FALSE}
# Load necessary libraries
library(googlesheets4)
library(readr)
library(tidyverse)
library(janitor)
library(lubridate)
library(kableExtra)
load("data/data.RData")
```

## Cleaning Script

During the time that the students were assembling data, we worked out of a Google Sheet as this permitted each of us to have simultaneous access to the data. For preliminary analysis, a data frame is constructed by reading from the sheet.

```{r message=FALSE, warning=FALSE, echo=TRUE}

# Read in the sheet
df_inspect <- read_sheet("https://docs.google.com/spreadsheets/d/1im5VSi3bIEi13O8WQ56wEIXSyNEstbGMylXXgD9bAG0/edit#gid=1858227071",
                 sheet="Inspections",
                 col_names = TRUE,
                 col_types = "c") %>% 
  clean_names()
```

Sometimes facility names are repeated. Use of the `dplyr` function `distinct` is useful for isolating issues with field names [@R-dplyr].

```{r df-inspect-before, message=FALSE, warning=FALSE, echo=FALSE}
df_inspect %>% 
  select(id, facility) %>% 
  distinct(facility, .keep_all = TRUE) %>% 
  arrange(facility) %>% 
  
  # Do some styling to the table
  kable(caption = "Facility Names Before Cleaning",
        col.names = c("ID", "Facility")) %>% 
  kable_paper(bootstrap_options = c(
    "stripped",
    "hover",
    "condensed",
    "responsive"
  )) %>% 
  scroll_box(height="400px")
```

The following procedures detail the main facility name cleaning operations. This block of code was converted to a custom function and written to an R file. By doing this, the entire following code block can be called with a single function, or incorporated into an analysis specific pipeline. The cleaning procedures are provided here to show the transformations, but the separate function file can be [downloaded](https://github.com/n8craig/inspections/blob/main/function_clean_facility_names.R) directly. Heavy use is made of the `dplyr` library [@R-dplyr] `mutate` function in combination with the `stringr` library [@R-stringr] functions `str_replace_all` and `replace` functions. All of these are part of the `tidyverse` [@R-tidyverse] family of libraries.

```{r df-clean-names, message=FALSE, warning=FALSE, echo=TRUE}
df_inspect <- df_inspect %>% 
  # Removing a curious whitespace character that appears
  # in more recent facility names
  mutate(facility = str_replace_all(facility,
         pattern = " ",
         replacement = " ")) %>% 
  
  separate(.,
           col = facility,
           into = c("facility","inspection_date"),
           sep = "\\) - ") %>%
  separate(.,
           col = facility,
           into = c("facility","state"),
           sep = "\\(") %>% 

  # Deal with issues in the file names, state, and inspection date
  mutate(facility = str_trim(facility, side = "both"),
         state = str_trim(state, side = "both"),
         facility = str_replace_all(facility,
                            pattern = "^Adelanto ICE Processing Center-East",
                            replacement = "Adelanto ICE Processing Center - East"),
         facility = str_replace_all(facility,
                            pattern = "^Adelanto ICE Processing Center-West",
                            replacement = "Adelanto ICE Processing Center - West"),
         facility = str_replace_all(facility,
                            pattern = "^Allen Parish Detention Facility",
                            replacement = "Allen Parish Public Safety Complex"),
         facility = str_replace_all(facility,
                            pattern = "^Berks County Residential Center",
                            replacement = "Berks Family Residential Center"),
         inspection_date = replace(inspection_date, inspection_date=="Jan. 29 - 31, 2019", "Jan. 31, 2019"),
         facility = str_replace_all(facility,
                            pattern = "^Bristol County Jail$",
                            replacement = "Bristol County Jail and House of Correction"),
         facility = str_replace_all(facility,
                            pattern = "^Buffalo$",
                            replacement = "Buffalo Batavia Service Processing Center"),
         state = replace(state, state == "Batavia) Service Processing Center", "NY"),
         facility = str_replace_all(facility,
                            pattern = "^Calhoun County Jail",
                            replacement = "Calhoun County Correctional Center"),
         facility = str_replace_all(facility,
                            pattern = "^Clay County Justice Center",
                            replacement = "Clay County Jail"),
         facility = str_replace_all(facility,
                            pattern = "^Coastal Bend Detention Facility",
                            replacement = "Coastal Bend Detention Center"),
         state = replace(state, state == "David L. Moss Criminal Justice Center)", "OK"),
         facility = str_replace_all(facility,
                            pattern = "^Dodge County Detention Center",
                            replacement = "Dodge County Detention Facility"),
         facility = str_replace_all(facility,
                            pattern = "^Donald W. Wyatt Detention Center",
                            replacement = "Donald W. Wyatt Detention Facility"),
         facility = str_replace_all(facility,
                            pattern = "^Essex County Corrections Facility",
                            replacement = "Essex County Correctional Facility"),
         facility = str_replace_all(facility,
                            pattern = "^Farmville Detention Center$",
                            replacement = "Immigration Centers of America - Farmville"),
         facility = str_replace_all(facility,
                            pattern = "^Florence SPC",
                            replacement = "Florence Service Processing Center"),
         facility = str_replace_all(facility,
                            pattern = "^Houston CDF",
                            replacement = "Houston Contract Detention Facility"),
         state = replace(state, state == "Polk)", "TX"),
         facility = str_replace_all(facility,
                            pattern = "^Immigration Centers of America$",
                            replacement = "Immigration Centers of America - Farmville"),
         state = replace(state, state =="ICA", "VA"),
         inspection_date = replace(inspection_date,
                                   inspection_date == "Farmville Detention Center (FDC) (VA",
                                   "Feb. 24, 2021"),
         # This one picks a value in the facility col and changes a value in the state col
         state = replace(state, facility == "Immigration Centers of America - Farmville",
                         "VA"),
         facility = str_replace_all(facility,
                            pattern = "^Joe Corley Detention Facility",
                            replacement = "Joe Corley Processing Center"),
         facility = str_replace_all(facility,
                            pattern = "^Karnes County Residential Center",
                            replacement = "Karnes County Family Residential Center"),
         facility = str_replace_all(facility,
                            pattern = "^Krome SPC",
                            replacement = "Krome Service Procesing Center"),
         facility = str_replace_all(facility,
                            pattern = "^Krome Special Processing Center",
                            replacement = "Krome Service Procesing Center"),
         state = replace(state, state =="SPC)", "FL"),
         facility = str_replace_all(facility,
                            pattern = "^Mesa Verde Detention Facility",
                            replacement = "Mesa Verde ICE Processing Facility"),
         facility = str_replace_all(facility,
                            pattern = "^Northwest Contract Detention Center",
                            replacement = "Northwest ICE Processing Center"),
         facility = str_replace_all(facility,
                            pattern = "^Northwest Detention Center",
                            replacement = "Northwest ICE Processing Center"),
         facility = str_replace_all(facility,
                            pattern = "^Okmulgee County Jail-Moore Detention Facility",
                            replacement = "Okmulgee County Jail - Moore Detention Facility"),
         facility = str_replace_all(facility,
                            pattern = "^Orange County Jail",
                            replacement = "Orange County Correctional Facility"),
         facility = str_replace_all(facility,
                            pattern = "^Otay Mesa Detention Facility",
                            replacement = "Otay Mesa Detention Center"),
         facility = str_replace_all(facility,
                            pattern = "^Prarieland Detention Center",
                            replacement = "Prairieland Detention Center"),
         facility = str_replace_all(facility,
                            pattern = "^Richwood Correcrtional Center",
                            replacement = "Richwood Correctional Center"),
         facility = str_replace_all(facility,
                            pattern = "^Rio Grande Processing Center",
                            replacement = "Rio Grande Detention Center"),
         facility = str_replace_all(facility,
                            pattern = "^Robert A. Deyton Correctional Center",
                            replacement = "Robert A. Deyton Detention Facility"),
         facility = str_replace_all(facility,
                            pattern = "^Robert A. Deyton Detention Center",
                            replacement = "Robert A. Deyton Detention Facility"),
         facility = str_replace_all(facility,
                            pattern = "^South Texas Detention Complex",
                            replacement = "South Texas ICE Processing Center"),
         facility = str_replace_all(facility,
                            pattern = "^South Texas Processing Center",
                            replacement = "South Texas ICE Processing Center"),
         facility = str_replace_all(facility,
                            pattern = "^Strafford County Corrections",
                            replacement = "Strafford County Department of Corrections"),
         inspection_date = replace(inspection_date, state =="CO)- Mar. 31, 2021", "Mar. 31, 2021"),
         state = replace(state, state == "CO)- Mar. 31, 2021", "CO"),
         facility = str_replace_all(facility,
                            pattern = "^T. Don Hutto Detention Center$",
                            replacement = "T. Don Hutto Residential Center"),
         facility = str_replace_all(facility,
                            pattern = "Tulsa County Jail - David L. Moss Criminal Jutice Center",
                            replacement = "David L. Moss Criminal Justice Center"),
         facility = str_replace_all(facility,
                            pattern = "^Tulsa County Jail$",
                            replacement = "David L. Moss Criminal Justice Center"),
         state = replace(state, state == "David L. Moss Justice Center)", "OK"),
         facility = str_replace_all(facility,
                            pattern = "^Washoe County Detention Centerr",
                            replacement = "Washoe County Detention Center"),
         facility = str_replace_all(facility,
                            pattern = "^Webb County Detention Facility",
                            replacement = "Webb County Detention Center"),
         )
```

```{r df-inspect-after}
df_inspect %>% 
  select(id, facility) %>% 
  distinct(facility, .keep_all = TRUE) %>% 
  arrange(facility) %>% 
  
  # Do some styling to the table
  kable(caption = "Facility Names After Cleaning",
        col.names = c("ID", "Facility")) %>% 
  kable_paper(bootstrap_options = c(
    "stripped",
    "hover",
    "condensed",
    "responsive"
  )) %>% 
  scroll_box(height="400px")
```

```{r package-references-appendix, echo=FALSE}
knitr::write_bib(file = 'packages.bib')
```

# Facet Plots

## Assaults

```{r assault-facet, fig.height=40}

# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(5, "Paired")
plot_colors[5] <- "#000000"

# Create the labels
plot_labels <- c(
    "Detained Person on Staff w. Serious Injury",
    "Detained Person on Staff w/o Serious Injury",
    "Detained Person on Detained Person w. Serous Injury",
    "Detained Person on Detained Person w/o Serous Injury",
    "Total Assaults")

df_assaults %>% 
  # Calling the plot and formatting
  ggplot(aes(x=date, y = assault_count, linetype=assault_type))+
  geom_line(aes(color = assault_type), size =.65) +
  
  # Set the line type
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Assault Type:",
    guide = guide_legend(nrow = 5)
    )+
  
  # Setting the color
  scale_color_manual(
    values = plot_colors,
    labels = plot_labels,
  name = "Assault Type:"
  )+

  labs(title = "Reported Assaults")+
  ylab("Number of Assaults")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =5))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

## Disciplinary

```{r disciplinary-facet, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "solid",
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(6, "Paired")
plot_colors[6] <- "#000000"

# Create plot labels

plot_labels <- c(
    "Disciplinary Infractions",
    "Disciplinary Infractions - Guilty",
    "Disciplinary Appeals",
    "Disciplinary Appeals in Favor of Detainee",
    "Sanctions Over 60 Days",
    "Total Disciplinary")


df_discipline %>% 
  
  # Calling the plot and formatting
  ggplot(aes(x=date,
             y = disciplinary_count,
             linetype = disciplinary_type))+
  geom_line(aes(color = disciplinary_type), size = .65) +
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Disciplinary Type:",
    guide = guide_legend(nrow = 6)
    )+
  
  # Setting the color
  scale_color_manual(
  values = plot_colors,
  labels = plot_labels,
  name = "Disciplinary Type:"
  )+

  labs(title = "Disciplinary")+
  ylab("Number of Disciplinary Infractions")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =6))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

## Solitary

```{r solitary-facet, fig.height=40, fig.cap="Facet plot of solitary instances reported by facility over time"}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(5, "Paired")
plot_colors[5] <- "#000000"

# Create plot labels
plot_labels <- c(
    "Administrative",
    "Disciplinary",
    "Medical",
    "Mental Health",
    "Total")

df_solitary %>% 

# Calling the plot and formatting
  ggplot(aes(x=date,
             y = segregation_count,
             linetype = segregation_type))+
  geom_line(aes(color = segregation_type), size = .65) +
  
  # Set the color
  scale_color_manual(
  values = plot_colors,
  name = "Solitary Type:",
  labels = plot_labels)+
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
      name = "Solitary Type:",
    labels = plot_labels)+
  
  
  labs(title = "Reported Use of Solitary")+
  ylab("Number of Individuals Palced in Solitary")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+
    facet_wrap(~ facility, ncol=3)
```

## Use of Force

```{r use-force-facet, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "solid",
    "solid",
    "solid",
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(8, "Paired")
plot_colors[8] <- "#000000"

# Create plot labels
plot_labels <- c(
    "Immediate Use of Force Incidents",
    "Calculated Use of Force Incidents",
    "Uses of Force with Chemical Agents",
    "Incidents Where Non-Lethal Weapons Were Used",
    "Times 4/5 Point Restraints Used",
    "Use of Force With Serious Injury",
    "Strip Searches",
    "Total Disciplinary")

df_force %>% 
  
  # Calling the plot and formatting
  ggplot(aes(x=date, y = use_of_force_count, linetype=use_of_force_type))+
  geom_line(aes(color = use_of_force_type), size = .65) +
  
  # setting the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Use of Force Type:",
    guide = guide_legend(nrow = 8)
    )+
  
  # Setting the color
  scale_color_manual(
  values = plot_colors,
  labels = plot_labels,
  name = "Use of Force Type:",
  )+

  labs(title = "Disciplinary")+
  ylab("Number of Disciplinary Infractions")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

    # Set the legend to multiple rows
  guides(col = guide_legend(nrow =8))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

## Sexual Abuse and Assault

### Allegations

```{r sexual-abuse-allegation-facet, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "solid",
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(6, "Paired")
plot_colors[6] <- "#000000"

# Create plot labels
plot_labels <- c(
    "Sexual Abuse Allegations - Detainee on Detainee",
    "Sexual Abuse Allegations - Inmate on Detainee",
    "Sexual Abuse Allegations - Detainee on Inmate",
    "Sexual Abuse Allegations - Staff/Contractor/Volunteer on Detainee",
    "Number of Sexual Abuse Allegations - Detainee on Staff/Contractor/Volunteer",
    "Total Sexual Abuse Allegations")

df_sex_alleg %>% 
  # Calling the plot and formatting
  ggplot(aes(x=date, y = sexual_abuse_allegations_count, linetype=sexual_abuse_allegations_type))+
  geom_line(aes(color = sexual_abuse_allegations_type),size=.65) +
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Sexual Abuse Allegation Type:",
    guide = guide_legend(nrow = 6)
    )+
  
  # Set the color
  scale_color_manual(values = plot_colors,
                     labels = plot_labels,
                     name = "Sexual Abuse Allegation Type:")+

  labs(title = "Sexual Abuse Allegations")+
  ylab("Number of Allegations")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

    # Set the legend to multiple rows
  guides(col = guide_legend(nrow =8))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

### Allegations Substantiated

```{r sexual-abuse-substantiated-facet, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "solid",
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(6, "Paired")
plot_colors[6] <- "#000000"

# Create the labels
plot_labels <- c(
    "Sexual Abuse Allegations - Detainee on Detainee",
    "Sexual Abuse Allegations - Inmate on Detainee",
    "Sexual Abuse Allegations - Detainee on Inmate",
    "Sexual Abuse Allegations - Staff/Contractor/Volunteer on Detainee",
    "Number of Sexual Abuse Allegations - Detainee on Staff/Contractor/Volunteer",
    "Total Sexual Abuse Allegations")

df_sex_alleg_sub %>% 
  
  # Calling the plot and formatting
  ggplot(aes(x=date, y = sexual_abuse_substantiated_count, linetype=sexual_abuse_substantiated_type))+
  geom_line(aes(color = sexual_abuse_substantiated_type),size=.65) +
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Sexual Abuse Allegation Type:",
    guide = guide_legend(nrow = 6)
    )+
  
  # Set the color
  scale_color_manual(values = plot_colors,
                     labels = plot_labels,
                     name = "Sexual Abuse Allegation Type:")+

  labs(title = "Sexual Abuse Allegations Substantiated")+
  ylab("Number of Allegations")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =8))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

## Medical

### Medical Mental Observation

```{r medical-mental-observation-facet, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors
plot_colors <- brewer.pal(3, "Paired")
plot_colors[3] <- "#000000"

# Create plot labels
plot_labels <- c(
    "Detainees in Medical Observation",
    "Detainees in Mental Health Observation",
    "Total Individuals in Medical or Mental Health Observation")

df_observation %>% 

  # Calling the plot and formatting
  ggplot(aes(x=date, y = medical_mental_observation_count, linetype=medical_mental_observation_type))+
  geom_line(aes(color = medical_mental_observation_type),size=.65) +
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Observation Type:",
    guide = guide_legend(nrow = 3)
    )+
  
  # Set the color
  scale_color_manual(values = plot_colors,
                     labels = plot_labels,
                     name = "Observation Type:",)+

  labs(title = "Medical and Mental Health Observation")+
  ylab("Number of Individuals in Observation")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =3))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

### Infectious Disease

```{r infectious-disease-facet, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid"
    )

# Use Color Brewer to set colors
plot_colors <- brewer.pal(2, "Paired")

# Create plot labels
plot_labels <- c(
    "Infectious Disease Reported",
    "Infectious Disease Confirmed")

df_infectious %>% 
  
  # Calling the plot and formatting
  ggplot(aes(x=date, y = infectious_disease_count, linetype=infectious_disease_type))+
  geom_line(aes(color = infectious_disease_type),size=.65) +
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Category:",
    guide = guide_legend(nrow = 2)
    )+
  
  # Set the color
  scale_color_manual(values = plot_colors,
    labels = plot_labels,
    name = "Category:")+

  labs(title = "Infectious Diseases Reported and Confirmed")+
  ylab("Count")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =3))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

### Medical and Mental Health Referrals

```{r referrals-facet, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors
plot_colors <- brewer.pal(5, "Paired")
plot_colors[5] <- "#000000"

# Create plot labels
plot_labels <- c(
    "Outside Medical Referrals",
    "Detainees Transported to Off-Site Hospitals for Emergency Reasons",
    "Admissions to Off-Site Hospitals for Medical Reasons",
    "Admissions to Off-Site Hospitals for Mental Health Reasons",
    "Total Referrals")

df_referrals %>% 
  # Calling the plot and formatting
  ggplot(aes(x=date, y = referral_count, linetype=referral_type))+
  geom_line(aes(color = referral_type),size=.65) +
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Referral Type:",
    guide = guide_legend(nrow = 5)
    )+
  
  # Set the color
  scale_color_manual(values = plot_colors,
    labels = plot_labels,
    name = "Referral Type:")+

  labs(title = "Medical and Mental Health Referrals")+
  ylab("Number of Referrals")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =5))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

### Sick Call Requests and Encounters

```{r sick-calls-plot, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors
plot_colors <- brewer.pal(3, "Paired")
plot_colors[3] <- "#000000"

# Create labels
plot_labels <- c(
    "Sick Call Requests",
    "Sick Call Encounters",
    "Total Sick Calls")

df_324_inc %>%
  
  # Subset the df to only the used cols
  select(id, facility, date,
       sick_call_requests:
       sick_call_encounters
       ) %>% 
  
  # Need the rowwise function to compute a row-at-a-time
  # in the following mutate function
  rowwise(id) %>% 
  
  # Create a new total column
  mutate(total_sick_calls = sum(c_across(
       sick_call_requests:
       sick_call_encounters
  ), na.rm = TRUE)) %>% 

  # Call a range of table columns and pivot long
  pivot_longer(.,
               cols=  sick_call_requests:total_sick_calls,
               names_to = "sick_call_type",
               values_to = "sick_call_count") %>%
  
  # Remove NA values
  drop_na() %>% 
  
  # Explicitly define factor levels
  mutate(sick_call_type = factor(sick_call_type, levels = c(
    "sick_call_requests",
    "sick_call_encounters",
    "total_sick_calls"
  ))) %>% 
  
  # Calling the plot and formatting
  ggplot(aes(x=date, y = sick_call_count, linetype=sick_call_type))+
  geom_line(aes(color = sick_call_type),size=.65) +
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Sick Call Type:",
    guide = guide_legend(nrow = 3)
    )+
  
  # Set the color
  scale_color_manual(values = plot_colors,
    labels = plot_labels,
    name = "Sick Call Type:")+

  labs(title = "Sick Calls")+
  ylab("Number of Sick Calls")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =3))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

### Suicide Attempts and Watches

```{r suicide-facet, fig.height=40}

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(2, "Paired")

# Create plot labels
plot_labels <- c(
      "Suicide Attempts or Self Harm",
      "Suicide Watches/Constant Watch/Mental Health Observation")

# Calling the plot and formatting
df_suicide %>% 
  ggplot(aes(x=date, y = suicide_count))+
  geom_line(aes(color=suicide_type), size = .65) +
  
  # Set the color
  scale_color_manual(
    values = plot_colors,
    labels = plot_labels,
    name = "Type:"
    )+
  
  labs(title = "Suicide Attempts and Watches")+
  ylab("Number of Attempts or Watches")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 5),
    legend.position = "bottom"
    )+
  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =5))+
  
  # Wrap
  facet_wrap(~ facility, ncol = 3)
```

### Hunger Strikes

```{r hunger-strike-facet, fig.height=40, fig.cap="Facet plot of hunger strikes per month by facility."}
# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(2, "Paired")

df_hunger_strike %>% 
  
  # Calling the plot and formatting
  ggplot(aes(x=date, y = hunger_strikes))+
  geom_line(size = .65) +
  

  labs(title = "Hunger Strikes")+
  ylab("Number of Hunger Strikes")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 5),
    legend.position = "bottom"
    )+
  
  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =5))+
  
  # Wrap
  facet_wrap(~ facility, ncol = 3)
```

## Cause of Death
