---
title: "Preliminary Reading and Graphing of Inspection Data"
author: "Nathan Craig"
date: "`r format(Sys.Date(), '%A %B %d, %Y')`"
output:
  bookdown::html_document2: 
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
    code_download: true
bibliography: [citations.bib, packages.bib]
---

During the Summer of 2021, data from ICE facility inspections was compiled into a series of spreadsheets. The impetus for the project was Nathan Craig who was interested in collating the data to examine them for trends, patterns, and anomalies. The work was performed by three students funded under a National Science Foundation grant awarded to Dr. Neil Harvey (NMSU) and Dr. Jeremy Slack (NMSU). The students who performed the work were Juan Becerra (Stanford), Avigail Turima Romo (Columbia), and Daniela Navarro Verdugo (CSSLO). The three students worked with Nathan Craig (NMSU and AVID) to develop a workflow for compiling individual report significant incident summary forms and inspection cover letters into a series of spreadsheets.

This document is a preliminary proof of concept and initial exploratory data analysis [@tukey1977]. Goals of the document are to:

-   illustrate the kinds of analysis possible with the tabulation of ICE inspection data
-   demonstrate code for reading the spreadsheet
-   show specific wrangling methods unique to the data structure
-   provide preliminary exploratory plot of select variables from the Significant Incident Summary Sheets.

The data are presently incomplete and represent results compiled as of `r Sys.Date()`.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r load-libraries}
# Load necessary libraries

# Reading and wrangling
library(googlesheets4)
library(readr)
library(tidyverse)
library(janitor)
library(lubridate)
library(DT)

# Plotting
library(ggplot2)
library(RColorBrewer)

# Tables
library(kableExtra)
```

# Introduction

## Method for Reading G-324A-19-inc form

The code for this report is specific to data logged in the G-324A-19 form and the incident summaries portion of that form. The incident summary portion of this form is probably the most complex data structure in the project and also contains the largest volume of quantitative information. Therefore, it seemed like a good place to start with an initial proof of concept.

At the outset, there were issues reading in the Google Sheet. Several of the numeric columns read into R as lists which was undesirable. In the Google Sheet, Craig set the columns to `plain text` rather than `auto` and the data read in fine as character columns. This seemed to resolve the issue for a time, and it was possible to use the `type_convert()` function from the `readr` library [@R-readr] to convert these columns to numbers. Additional steps were taken to use the `read_sheet` function of the `googlesheets4` library [@R-googlesheets4] to explicitly set the column types to character.

There were also some issues using `lubridate` [@R-lubridate] to wrangle the month and year fields to a proper date. Craig was not able to discern the cause of the issue. However, changing the month format on the incident sheet from abbreviated to fully written out months solved the issue. Those changes were made in the master Google Sheet and the data were called using the `googlesheets4` library.

```{r read-data-324}
# Read in Sheet G-324A-19
df_324 <- read_sheet("https://docs.google.com/spreadsheets/d/1im5VSi3bIEi13O8WQ56wEIXSyNEstbGMylXXgD9bAG0/edit#gid=1858227071",
                 sheet="G-324A-19",
                 col_names = TRUE,
                 col_types = "c") %>% 
  clean_names() %>% 

  # Removing a curious whitespace character that appears
  # in more recent facility names
  mutate(facility = str_replace_all(facility,
         pattern = " ",
         replacement = " ")) %>% 
  
  
  # Clean up the facility names separating out inspection date and state
  # Note the use of a double escape character. This was necessary.
  separate(.,
           col = facility,
           into = c("facility","inspection_date"),
           sep = "\\) - ") %>%
  separate(.,
           col = facility,
           into = c("facility","state"),
           sep = "\\(") %>% 

  # Convert the character type cols to numbers
  # type_convert() %>%
  mutate(facility = as.factor(facility),
         state = as.factor(state),
         date = mdy(inspection_date),
         current_inspection_date_from = mdy(current_inspection_date_from),
         current_inspection_date_to = mdy(current_inspection_date_to)
         ) %>% 
  relocate(date, .before = inspection_date) %>% 
  mutate_at(c(20:49), as.numeric)
```




```{r read-data-324-incident}
# Read Google Sheet incident worksheet, convert to data frame, and wrangle
df_324_inc <- read_sheet("https://docs.google.com/spreadsheets/d/1im5VSi3bIEi13O8WQ56wEIXSyNEstbGMylXXgD9bAG0/edit#gid=1858227071",
                 sheet="G-324A-19-inc",
                 col_types = "c") %>% 
  clean_names() %>% 

  # Clean up the facility names separating out inspection date and state
  # Note the use of a double escape character. This was necessary.
  separate(.,
           col = facility,
           into = c("facility","inspection_date"),
           sep = "\\) - ") %>%
  separate(.,
           col = facility,
           into = c("facility","state"),
           sep = "\\(") %>% 

  # Convert the character type cols to numbers
  # type_convert() %>% 
  unite(date, year:month) %>% 
  mutate(facility = as.factor(facility),
         state = as.factor(state),
         date = ym(date)
         ) %>% 
  mutate_at(c(6:76), as.numeric)
```

At present `r nrow(df_324)` out of approximately 300 are complete. The inspections range in time from `r min(df_324$date, na.rm = TRUE)` to `r max(df_324$date, na.rm = TRUE)`.

## Summary Tables

Summary tables help to provide an overview of how many instances occur within particular category of data or how frequent a particular issue is recorded at a given facility. Summary tables are generated using a combination of `group_by` with `summarize` with the results piped to `kable` and `kableExtra`.

```{r inspection-count}
df_facility <- df_324 %>% 
  group_by(facility) %>% 
  summarise(n_inspections = n()) %>% 
  ungroup() 
  
df_facility %>% 
  kable(caption = "Inspections Reviewed by Facility",
        col.names = c("Facility", "Inspections Reviewed")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "300px")
```

## Facet Plots

Facet plots are produced for several categories of data. Within each category, columns are pivoted longer and plotted by date. Graphing is done using the `ggplot2` library [@R-ggplot2] with the `facet_wrap` function to provide a means to compare multiple facilities simultaneously. Such a plot can help identify trends and guide more specific questions.

# Assaults

The ICE G-324A Significant Incident Summary sheet contains information on a number of different kinds of assaults. Specifically, there are four categories of assaults listed:

-   Detained Person on Staff with Serious Injury
-   Detained Person on Staff without Serious Injury
-   Detained Person on Detained Person with Serious Injury
-   Detained Person on Detained Person without Serious Injury

Conspicuously absent from this list of assault types are assaults on detained persons by staff. This is deeply concerning because already initial review of inspection forms and prior news stories indicates that staff assaults on detained persons do occur. For example, at the Bristol County County Jail and House of Corrections independent investigation by the Massachusetts Office of Attorney General (MAOG) found that facility staff found a "planned and deliberate-use of force against the ICE detainees that was disproportionate to the security needs at that time and unnecessarily caused, or risked causing, harm to all involved" [@maag2020, ii]. The MOAG found that facility staff used "a flash bang grenade, pepper-ball launchers, pepper spray canisters, anti-riot shields, and canines--against detainees who had exhibited calm and nonviolent behavior for at least an hour before this operation" [@maag2020, ii]. Investigations found that staff "deployed these weapons both indiscriminately upon entry and also specifically against detainees who were not combative, assaultive, or otherwise actively resisting staff" [@maag2020, ii].

While the aforementioned Bristol County Jail may be a particularly egregious instance of staff assaulting detained persons, it is not an isolated case. For example, [video evidence](https://www.youtube.com/watch?v=JP_PfOhdLuM) from the Torrance County Detention Facility in Estancia NM shows CoreCivic's private facility staff pepper spraying persons detained by ICE on hunger strike [@swetlitz2020]. Review of inspections has also revealed additional inappropriate use of force incidents. The omission of a field to record instances of staff assaulting detained persons occludes and obscures real instances of violence. As the inspections process is a detention reform designed to make a more humane system of civil incarceration, the omission of categories for logging staff physical mistreatment of detained persons raises serious questions about the inspections processes ability to record significant and documented civil and human rights violations.

## Summary Tables

```{r assault-df}
df_assaults <- df_324_inc %>%
  
  # Subset the df to only the used cols
  select(id, facility, date,
       detainee_physical_assault_on_staff_with_serious_injury:
       detainee_on_detainee_physical_assault_fight_with_no_serious_injury
       ) %>% 
  
  # Need the rowwise function to compute a row-at-a-time
  # in the following mutate function
  rowwise(id) %>% 
  
  # Create a new total column
  mutate(total_assaults = sum(c_across(
       detainee_physical_assault_on_staff_with_serious_injury:
       detainee_on_detainee_physical_assault_fight_with_no_serious_injury
  ))) %>% 

  # Call a range of table columns and pivot long
  pivot_longer(.,
               cols=  detainee_physical_assault_on_staff_with_serious_injury:total_assaults,
               names_to = "assault_type",
               values_to = "assault_count") %>% 
  
  # Remove NA Values
  drop_na() %>% 
  
  # Explicitly define factor levels
  mutate(assault_type = factor(assault_type, levels = c(
    "detainee_physical_assault_on_staff_with_serious_injury",
    "detainee_physical_assault_on_staff_with_no_serious_injury",
    "detainee_on_detainee_physical_assault_fight_with_serious_injury",
    "detainee_on_detainee_physical_assault_fight_with_no_serious_injury",
    "total_assaults"
  )))
```

```{r assault-summary}
df_assaults %>% 
  group_by(assault_type) %>% 
  summarise(`Total Assaults by Type` = sum(assault_count)) %>% 
  ungroup() %>% 
  kable(caption = "Total Assaults by Type",
        col.names = c("Assault Type", "Total Assault Type")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))


df_assaults %>% 
  group_by(facility) %>% 
  summarise(total_assaults = sum(assault_count)) %>% 
  arrange(desc(total_assaults)) %>% 
  ungroup() %>% 
  kable(caption = "Total Assaults by Facility",
        col.names = c("Facility", "Total Assaults by Facility")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "300px")
```

## Facet of Assaults by Plot

The figure below is a facet plot of the assault categories present on the G-324A SIS form.

```{r assault-facet, fig.height=40}

# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(5, "Paired")
plot_colors[5] <- "#000000"

# Create the labels
plot_labels <- c(
    "Detained Person on Staff w. Serious Injury",
    "Detained Person on Staff w/o Serious Injury",
    "Detained Person on Detained Person w. Serous Injury",
    "Detained Person on Detained Person w/o Serous Injury",
    "Total Assaults")

df_assaults %>% 
  # Calling the plot and formatting
  ggplot(aes(x=date, y = assault_count, linetype=assault_type))+
  geom_line(aes(color = assault_type), size =.65) +
  
  # Set the line type
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Assault Type:",
    guide = guide_legend(nrow = 5)
    )+
  
  # Setting the color
  scale_color_manual(
    values = plot_colors,
    labels = plot_labels,
  name = "Assault Type:"
  )+

  labs(title = "Reported Assaults")+
  ylab("Number of Assaults")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =5))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

# Disciplinary

As a matter of law, ICE detention is not punitive it is purely administrative civil holding to ensure that individuals are present for their immigration proceedings. Despite that fact, there is a tremendous amount of punitive disciplinary measures that take place in ICE detention. Some of this information on discipline in ICE detention is logged in facility inspections. As seen in the tables below, during the current sample period there were more than 42,000 disciplinary infractions.

## Summary Tables

```{r}
df_discipline <- df_324_inc %>%
  
  # Subset the df to only the used cols
  select(id, facility, date,
       disciplinary_infractions:
       sanctions_over_60_days
       ) %>% 
  
  # Need the rowwise function to compute a row-at-a-time
  # in the following mutate function
  rowwise(id) %>% 
  
  # Create a new total column
  mutate(total_disciplinary = sum(c_across(
       disciplinary_infractions:
       sanctions_over_60_days
  ))) %>% 

  # Call a range of table columns and pivot long
  pivot_longer(.,
               cols=  disciplinary_infractions:total_disciplinary,
               names_to = "disciplinary_type",
               values_to = "disciplinary_count") %>%
  
  # Remove NA values
  drop_na() %>% 
  
  # Explicitly set factor levels
  mutate(disciplinary_type = factor(disciplinary_type, levels =c(
    "disciplinary_infractions",
    "disciplinary_infractions_guilty",
    "disciplinary_appeals",
    "disciplinary_appeals_found_in_favor_of_detainee",
    "sanctions_over_60_days",
    "total_disciplinary"
  )))
```

```{r}
df_discipline %>% 
  group_by(disciplinary_type) %>% 
  summarise(`Total Disciplinary by Type` = sum(disciplinary_count)) %>% 
  ungroup() %>% 
  kable(caption = "Total Disciplinary by Type",
        col.names = c("Discipline Type", "Total Discipline Type")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))


df_discipline %>% 
  group_by(facility) %>% 
  summarise(total_disciplinary = sum(disciplinary_count)) %>% 
  arrange(desc(total_disciplinary)) %>% 
  ungroup() %>% 
  kable(caption = "Total Discipline by Facility",
        col.names = c("Facility", "Total Discipline by Facility")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "300px")
```

## Facet Plot of Disciplinary by Facility

```{r disciplinary-facet, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "solid",
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(6, "Paired")
plot_colors[6] <- "#000000"

# Create plot labels

plot_labels <- c(
    "Disciplinary Infractions",
    "Disciplinary Infractions - Guilty",
    "Disciplinary Appeals",
    "Disciplinary Appeals in Favor of Detainee",
    "Sanctions Over 60 Days",
    "Total Disciplinary")


df_discipline %>% 
  
  # Calling the plot and formatting
  ggplot(aes(x=date,
             y = disciplinary_count,
             linetype = disciplinary_type))+
  geom_line(aes(color = disciplinary_type), size = .65) +
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Disciplinary Type:",
    guide = guide_legend(nrow = 6)
    )+
  
  # Setting the color
  scale_color_manual(
  values = plot_colors,
  labels = plot_labels,
  name = "Disciplinary Type:"
  )+

  labs(title = "Disciplinary")+
  ylab("Number of Disciplinary Infractions")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =6))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

# Solitary

## Summary Tables

Of the present 163 inspections reviewed so far, there are more than 34,000 instances of solitary. That is roughly 208 instances of solitary per inspection.

```{r}
df_solitary <- df_324_inc %>%
  
  # Select a subset of columns to work with
  select(id,
         facility,
         date,
         detainees_placed_in_administrative_segregation:
           detainees_placed_in_segregation_for_mental_health_reasons) %>% 
  
  # Need the rowwise function to compute a row-at-a-time
  # in the following mutate function
  rowwise(id) %>% 
  
  # Create new total column
  mutate(total_segregation = sum(c_across(detainees_placed_in_administrative_segregation:
                 detainees_placed_in_segregation_for_mental_health_reasons))) %>% 
  
  # Tidy
  pivot_longer(.,
               cols= detainees_placed_in_administrative_segregation:
                 total_segregation,
               names_to = "segregation_type",
               values_to = "segregation_count") %>% 
  
  # Remove NA
  drop_na() %>% 
  
  # Explicitly set factor levels
  mutate(segregation_type = factor(segregation_type, levels = c(
    "detainees_placed_in_administrative_segregation",
    "detainees_placed_in_disciplinary_segregation",
    "detainees_placed_in_segregation_for_medical_reasons",
    "detainees_placed_in_segregation_for_mental_health_reasons",
    "total_segregation"
  )))
```

```{r}
df_solitary %>% 
  group_by(segregation_type) %>% 
  summarise(`Total Solitary by Type` = sum(segregation_count)) %>% 
  ungroup() %>% 
  kable(caption = "Total Solitary by Type",
        col.names = c("Solitary Type", "Total Solitary Type")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))


df_solitary %>% 
  group_by(facility) %>% 
  summarise(total_segregation = sum(segregation_count)) %>% 
  arrange(desc(total_segregation)) %>% 
  ungroup() %>% 
  kable(caption = "Total Solitary by Facility",
        col.names = c("Facility", "Total Solitary by Facility")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "300px")
```

## Facet Plots of Solitary by Facility

```{r solitary-facet, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(5, "Paired")
plot_colors[5] <- "#000000"

# Create plot labels
plot_labels <- c(
    "Administrative",
    "Disciplinary",
    "Medical",
    "Mental Health",
    "Total")

df_solitary %>% 

# Calling the plot and formatting
  ggplot(aes(x=date,
             y = segregation_count,
             linetype = segregation_type))+
  geom_line(aes(color = segregation_type), size = .65) +
  
  # Set the color
  scale_color_manual(
  values = plot_colors,
  name = "Solitary Type:",
  labels = plot_labels)+
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
      name = "Solitary Type:",
    labels = plot_labels)+
  
  
  labs(title = "Reported Use of Solitary")+
  ylab("Number of Individuals Palced in Solitary")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+
    facet_wrap(~ facility, ncol=3)
```

## Solitary Over 60 Days

```{r solitary-60-facet}

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(5, "Paired")
plot_colors[5] <- "#000000"

# Call the dataframe and select cols
df_324 %>% 
  select(id,
         facility,
         state,
         date,
         fac_operator,
         admin_seg_60_ice,
         disc_seg_60_ice) %>% 
  drop_na() %>% 
  
  # Need the rowwise function to compute a row-at-a-time
  # in the following mutate function
  rowwise(id) %>% 
  
  # Generate total col
  mutate(total_seg_60 = sum(c_across(
       admin_seg_60_ice:
       disc_seg_60_ice
  ))) %>%
  
  # Make tidy and filter
  pivot_longer(cols = admin_seg_60_ice:disc_seg_60_ice,
               names_to = "segregation_60_type",
               values_to = "segregation_60_count") %>%
  filter(segregation_60_type %in% c("admin_seg_60_ice", "disc_seg_60_ice")&
           segregation_60_count > 0) %>%
  
  # Initiate the plot and sort by sum
  ggplot(aes(x = segregation_60_count,
             y=reorder(fac_operator, segregation_60_count, sum),
             fill=segregation_60_type))+
  geom_bar(stat = "identity")+
  
  # Set the color fill
  scale_fill_brewer(type = "qual",
                    palette = "Paired",
                    name = "Segregation > 60 Type",
                    labels = c("Administrative",
                               "Disciplinary"))+
  labs(title= "Segregation > 60 Days by Facility Operator",
        x = "Segregation > 60 Days Count",
        y = "Facility Operator")+
  theme(legend.position = "bottom")
```

# Use of Force

## Summary Tables

```{r}
df_force <- df_324_inc %>%
  
  # Subset the df to only the used cols
  select(id, facility, date,
       immediate_use_of_force_incidents:
       strip_searches
       ) %>% 
  
  # Need the rowwise function to compute a row-at-a-time
  # in the following mutate function
  rowwise(id) %>% 
  
  # Create a new total column
  mutate(total_use_of_force = sum(c_across(
       immediate_use_of_force_incidents:
       strip_searches
  ))) %>% 

  # Call a range of table columns and pivot long
  pivot_longer(.,
               cols=  immediate_use_of_force_incidents:total_use_of_force,
               names_to = "use_of_force_type",
               values_to = "use_of_force_count") %>%
  
  # Remove NA values
  drop_na() %>% 
  
  # Explicitly define factors
  mutate(use_of_force_type = factor(use_of_force_type, levels = c(
    "immediate_use_of_force_incidents",
    "calculated_use_of_force_incidents",
    "uses_of_force_with_chemical_agents",
    "incidents_where_non_lethal_weapons_were_used",
    "number_of_times_4_5_point_restraints_were_used",
    "use_of_force_with_serious_injury",
    "strip_searches",
    "total_use_of_force"
  )))
```

```{r}
df_force %>% 
  group_by(use_of_force_type) %>% 
  summarise(`Total Use of Force by Type` = sum(use_of_force_count)) %>% 
  ungroup() %>% 
  kable(caption = "Total Use of Force by Type",
        col.names = c("Use of Force Type", "Total Use of Force Type")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))


df_force %>% 
  group_by(facility) %>% 
  summarise(total_use_of_force = sum(use_of_force_count)) %>% 
  arrange(desc(total_use_of_force)) %>% 
  ungroup() %>% 
  kable(caption = "Total Use of Force by Facility",
        col.names = c("Facility", "Total Use of Force by Facility")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "300px")
```

## Facet Plot of Use of Force by Facility

```{r use-force-facet, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "solid",
    "solid",
    "solid",
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(8, "Paired")
plot_colors[8] <- "#000000"

# Create plot labels
plot_labels <- c(
    "Immediate Use of Force Incidents",
    "Calculated Use of Force Incidents",
    "Uses of Force with Chemical Agents",
    "Incidents Where Non-Lethal Weapons Were Used",
    "Times 4/5 Point Restraints Used",
    "Use of Force With Serious Injury",
    "Strip Searches",
    "Total Disciplinary")

df_force %>% 
  
  # Calling the plot and formatting
  ggplot(aes(x=date, y = use_of_force_count, linetype=use_of_force_type))+
  geom_line(aes(color = use_of_force_type), size = .65) +
  
  # setting the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Use of Force Type:",
    guide = guide_legend(nrow = 8)
    )+
  
  # Setting the color
  scale_color_manual(
  values = plot_colors,
  labels = plot_labels,
  name = "Use of Force Type:",
  )+

  labs(title = "Disciplinary")+
  ylab("Number of Disciplinary Infractions")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

    # Set the legend to multiple rows
  guides(col = guide_legend(nrow =8))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

# Sexual Abuse and Assault

## Allegations

### Summary Tables

```{r}
df_sex_alleg <- df_324_inc %>%
  
  # Subset the df to only the used cols
  select(id, facility, date,
       sexual_abuse_allegations_detainee_on_detainee:
       sexual_abuse_allegations_detainee_on_staff_contractor_volunteer_29
       ) %>% 
  
  # Need the rowwise function to compute a row-at-a-time
  # in the following mutate function
  rowwise(id) %>% 
  
  # Create a new total column
  mutate(total_sexual_abuse_allegations = sum(c_across(
       sexual_abuse_allegations_detainee_on_detainee:
       sexual_abuse_allegations_detainee_on_staff_contractor_volunteer_29
  ))) %>% 

  # Call a range of table columns and pivot long
  pivot_longer(.,
               cols=  sexual_abuse_allegations_detainee_on_detainee:total_sexual_abuse_allegations,
               names_to = "sexual_abuse_allegations_type",
               values_to = "sexual_abuse_allegations_count") %>%
  
  # Remove NA
  drop_na() %>% 
  
  # Explicitly define factor levels
  mutate(sexual_abuse_allegations_type = factor(sexual_abuse_allegations_type, levels = c(
    "sexual_abuse_allegations_detainee_on_detainee",
    "sexual_abuse_allegations_inmate_on_detainee",
    "sexual_abuse_allegations_detainee_on_inmate",
    "sexual_abuse_allegations_detainee_on_staff_contractor_volunteer",
    "sexual_abuse_allegations_detainee_on_staff_contractor_volunteer_29",
    "total_sexual_abuse_allegations"
  )))
```

```{r}
df_sex_alleg %>% 
  group_by(sexual_abuse_allegations_type) %>% 
  summarise(`Sexual Abuse and Assault Allegations by Type` = sum(sexual_abuse_allegations_count)) %>% 
  ungroup() %>% 
  kable(caption = "Total Sexual Abuse and Assault Allegations by Type",
        col.names = c("Allegation Type", "Total Allegation Type")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))


df_sex_alleg %>% 
  group_by(facility) %>% 
  summarise(total_sexual_abuse_allegations = sum(sexual_abuse_allegations_count)) %>% 
  arrange(desc(total_sexual_abuse_allegations)) %>% 
  ungroup() %>% 
  kable(caption = "Total Sexual Abuse and Assault Allegations by Facility",
        col.names = c("Facility", "Total Allegations by Facility")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "300px")
```

### Facet Plot of Sexual Abuse and Assault Allegations by Facility

```{r sexual-abuse-allegation-facet, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "solid",
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(6, "Paired")
plot_colors[6] <- "#000000"

# Create plot labels
plot_labels <- c(
    "Sexual Abuse Allegations - Detainee on Detainee",
    "Sexual Abuse Allegations - Inmate on Detainee",
    "Sexual Abuse Allegations - Detainee on Inmate",
    "Sexual Abuse Allegations - Staff/Contractor/Volunteer on Detainee",
    "Number of Sexual Abuse Allegations - Detainee on Staff/Contractor/Volunteer",
    "Total Sexual Abuse Allegations")

df_sex_alleg %>% 
  # Calling the plot and formatting
  ggplot(aes(x=date, y = sexual_abuse_allegations_count, linetype=sexual_abuse_allegations_type))+
  geom_line(aes(color = sexual_abuse_allegations_type),size=.65) +
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Sexual Abuse Allegation Type:",
    guide = guide_legend(nrow = 6)
    )+
  
  # Set the color
  scale_color_manual(values = plot_colors,
                     labels = plot_labels,
                     name = "Sexual Abuse Allegation Type:")+

  labs(title = "Sexual Abuse Allegations")+
  ylab("Number of Allegations")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

    # Set the legend to multiple rows
  guides(col = guide_legend(nrow =8))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

## Substantiated Allegations

### Summary Tables

```{r}
df_sex_alleg_sub <- df_324_inc %>%
  
  # Subset the df to only the used cols
  select(id, facility, date,
       sexual_abuse_allegations_detainee_on_detainee_2:
       sexual_abuse_allegations_detainee_on_staff_contractor_volunteer_34
       ) %>% 
  
  # Need the rowwise function to compute a row-at-a-time
  # in the following mutate function
  rowwise(id) %>% 
  
  # Create a new total column
  mutate(total_sexual_abuse_allegations = sum(c_across(
       sexual_abuse_allegations_detainee_on_detainee_2:
       sexual_abuse_allegations_detainee_on_staff_contractor_volunteer_34
  ))) %>% 

  # Call a range of table columns and pivot long
  pivot_longer(.,
               cols=  sexual_abuse_allegations_detainee_on_detainee_2:total_sexual_abuse_allegations,
               names_to = "sexual_abuse_substantiated_type",
               values_to = "sexual_abuse_substantiated_count") %>%
  
  # Remove NA values
  drop_na() %>% 
  
  # Explicitly set factor levels
  mutate(sexual_abuse_substantiated_type= factor(sexual_abuse_substantiated_type, levels =c(
    "sexual_abuse_allegations_detainee_on_detainee_2",
    "sexual_abuse_allegations_inmate_on_detainee_2",
    "sexual_abuse_allegations_detainee_on_inmate_2",
    "sexual_abuse_allegations_staff_contractor_volunteer_on_detainee",
    "sexual_abuse_allegations_detainee_on_staff_contractor_volunteer_34",
    "total_sexual_abuse_allegations"
  )))
```

```{r}
df_sex_alleg_sub %>% 
  group_by(sexual_abuse_substantiated_type) %>% 
  summarise(`Sexual Abuse and Assault Allegations Substantiated by Type` = sum(sexual_abuse_substantiated_count)) %>% 
  ungroup() %>% 
  kable(caption = "Total Substantiated Sexual Abuse and Assault Allegations by Type",
        col.names = c("Substantiated Allegation Type", "Total Substantiated Allegation Type")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))


df_sex_alleg_sub %>% 
  group_by(facility) %>% 
  summarise(total_sexual_abuse_substantiated = sum(sexual_abuse_substantiated_count)) %>% 
  arrange(desc(total_sexual_abuse_substantiated)) %>% 
  ungroup() %>% 
  kable(caption = "Total Substantiated Sexual Abuse and Assault Allegations by Facility",
        col.names = c("Facility", "Total Substantiated Allegations by Facility")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "300px")
```

### Facet Plot of Substantiated Sexual Abuse and Assault Allegations

```{r sexual-abuse-substantiated-facet, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "solid",
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(6, "Paired")
plot_colors[6] <- "#000000"

# Create the labels
plot_labels <- c(
    "Sexual Abuse Allegations - Detainee on Detainee",
    "Sexual Abuse Allegations - Inmate on Detainee",
    "Sexual Abuse Allegations - Detainee on Inmate",
    "Sexual Abuse Allegations - Staff/Contractor/Volunteer on Detainee",
    "Number of Sexual Abuse Allegations - Detainee on Staff/Contractor/Volunteer",
    "Total Sexual Abuse Allegations")

df_sex_alleg_sub %>% 
  
  # Calling the plot and formatting
  ggplot(aes(x=date, y = sexual_abuse_substantiated_count, linetype=sexual_abuse_substantiated_type))+
  geom_line(aes(color = sexual_abuse_substantiated_type),size=.65) +
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Sexual Abuse Allegation Type:",
    guide = guide_legend(nrow = 6)
    )+
  
  # Set the color
  scale_color_manual(values = plot_colors,
                     labels = plot_labels,
                     name = "Sexual Abuse Allegation Type:")+

  labs(title = "Sexual Abuse Allegations Substantiated")+
  ylab("Number of Allegations")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =8))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

# Medical

## Medical Observation

```{r medical-mental-observation-facet, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors
plot_colors <- brewer.pal(3, "Paired")
plot_colors[3] <- "#000000"

# Create plot labels
plot_labels <- c(
    "Detainees in Medical Observation",
    "Detainees in Mental Health Observation",
    "Total Individuals in Medical or Mental Health Observation")

df_324_inc %>%
  
  # Subset the df to only the used cols
  select(id, facility, date,
       detainees_in_medical_observation:
       detainees_in_mental_health_observation
       ) %>% 
  
  # Need the rowwise function to compute a row-at-a-time
  # in the following mutate function
  rowwise(id) %>% 
  
  # Create a new total column
  mutate(total_mental_medical_observation = sum(c_across(
       detainees_in_medical_observation:
       detainees_in_mental_health_observation
  ))) %>% 

  # Call a range of table columns and pivot long
  pivot_longer(.,
               cols=  detainees_in_medical_observation:total_mental_medical_observation,
               names_to = "medical_mental_observation_type",
               values_to = "medical_mental_observation_count") %>%
  
  # Remove NA values
  drop_na() %>% 
  
  # Explicitly set factor levels
  mutate(medical_mental_observation_type = factor(medical_mental_observation_type, levels = c(
    "detainees_in_medical_observation",
    "detainees_in_mental_health_observation",
    "total_mental_medical_observation"
  ))) %>% 
  
  # Calling the plot and formatting
  ggplot(aes(x=date, y = medical_mental_observation_count, linetype=medical_mental_observation_type))+
  geom_line(aes(color = medical_mental_observation_type),size=.65) +
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Observation Type:",
    guide = guide_legend(nrow = 3)
    )+
  
  # Set the color
  scale_color_manual(values = plot_colors,
                     labels = plot_labels,
                     name = "Observation Type:",)+

  labs(title = "Medical and Mental Health Observation")+
  ylab("Number of Individuals in Observation")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =3))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
  
```

## Infections Disease

```{r infectious-disease-facet, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid"
    )

# Use Color Brewer to set colors
plot_colors <- brewer.pal(2, "Paired")

# Create plot labels
plot_labels <- c(
    "Infectious Disease Reported",
    "Infectious Disease Confirmed")

df_324_inc %>%
  
  # Subset the df to only the used cols
  select(id, facility, date,
       infectious_disease_reported:
       infections_disease_confirmed
       ) %>% 
  
  # Need the rowwise function to compute a row-at-a-time
  # in the following mutate function
  rowwise(id) %>% 
  
  # # Create a new total column
  # # In this case not used
  # mutate(total_infections_disease_report_confirmed = sum(c_across(
  #      infectious_disease_reported:
  #      infections_disease_confirmed
  # ))) %>% 

  # Call a range of table columns and pivot long
  pivot_longer(.,
               cols=  infectious_disease_reported:infections_disease_confirmed,
               names_to = "infectious_disease_type",
               values_to = "infectious_disease_count") %>%
  
  # Remove NA values
  drop_na() %>% 
  
  # Explicitly define factor levels
  mutate(infectious_disease_type = factor(infectious_disease_type, levels = c(
    "infectious_disease_reported",
    "infections_disease_confirmed"
  ))) %>% 
  
  # Calling the plot and formatting
  ggplot(aes(x=date, y = infectious_disease_count, linetype=infectious_disease_type))+
  geom_line(aes(color = infectious_disease_type),size=.65) +
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Category:",
    guide = guide_legend(nrow = 2)
    )+
  
  # Set the color
  scale_color_manual(values = plot_colors,
    labels = plot_labels,
    name = "Category:")+

  labs(title = "Infectious Diseases Reported and Confirmed")+
  ylab("Count")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =3))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
  
  
```

## Medical and Mental Health Referrals

```{r referrals-facet, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors
plot_colors <- brewer.pal(5, "Paired")
plot_colors[5] <- "#000000"

# Create plot labels
plot_labels <- c(
    "Outside Medical Referrals",
    "Detainees Transported to Off-Site Hospitals for Emergency Reasons",
    "Admissions to Off-Site Hospitals for Medical Reasons",
    "Admissions to Off-Site Hospitals for Mental Health Reasons",
    "Total Referrals")


df_324_inc %>%
  
  # Subset the df to only the used cols
  select(id, facility, date,
       outside_medical_referrals:
       admissions_to_off_site_hospitals_for_mental_health_reasons
       ) %>% 
  
  # Need the rowwise function to compute a row-at-a-time
  # in the following mutate function
  rowwise(id) %>% 
  
  # Create a new total column
  mutate(total_referrals = sum(c_across(
       outside_medical_referrals:
       admissions_to_off_site_hospitals_for_mental_health_reasons
  ))) %>% 

  # Call a range of table columns and pivot long
  pivot_longer(.,
               cols=  outside_medical_referrals:total_referrals,
               names_to = "referral_type",
               values_to = "referral_count") %>%
  
  # Remove NA values
  drop_na() %>% 
  
  # Explicitly define factor levels
  mutate(referral_type= factor(referral_type, levels = c(
    "outside_medical_referrals",
    "detainees_transported_to_off_site_hospitals_for_emergency_care",
    "admissions_to_off_site_hospitals_for_medical_reasons",
    "admissions_to_off_site_hospitals_for_mental_health_reasons",
    "total_referrals"
  ))) %>% 
  
  
  # Calling the plot and formatting
  ggplot(aes(x=date, y = referral_count, linetype=referral_type))+
  geom_line(aes(color = referral_type),size=.65) +
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Referral Type:",
    guide = guide_legend(nrow = 5)
    )+
  
  # Set the color
  scale_color_manual(values = plot_colors,
    labels = plot_labels,
    name = "Referral Type:")+

  labs(title = "Medical and Mental Health Referrals")+
  ylab("Number of Referrals")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =5))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

## Sick Call Requests and Encounters



```{r}
df_sick_call <- df_324_inc %>%
  
  # Select a subset of columns to work with
  select(id,
         facility,
         date,
         sick_call_requests,
         sick_call_encounters) %>% 
  
  # Need the rowwise function to compute a row-at-a-time
  # in the following mutate function
  rowwise(id) %>% 
  
  # Create new total column
  mutate(total_sick_call = sick_call_requests,
         sick_call_encounters) %>% 
  
  # Tidy
  pivot_longer(.,
               cols= sick_call_requests:
         sick_call_encounters,
               names_to = "sick_call_type",
               values_to = "sick_call_count") %>% 
  
  # Remove NA
  drop_na() %>% 
  
  # Explicitly set factor levels
  mutate(sick_call_type = factor(sick_call_type, levels = c(
    "sick_call_requests",
    "sick_call_encounters",
    "total_sick_call"
  )))
```

```{r}
df_sick_call %>% 
  group_by(sick_call_type) %>% 
  summarise(`Total Sick Call by Type` = sum(sick_call_count)) %>% 
  ungroup() %>% 
  kable(caption = "Total Sick Call by Type",
        col.names = c("Type", "Total")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))

df_sick_call %>% 
  group_by(facility) %>% 
  summarise(total_sick_call = sum(sick_call_count)) %>% 
  arrange(desc(total_sick_call)) %>% 
  ungroup() %>% 
  kable(caption = "Total Sick Call Requests/Encounters by Facility",
        col.names = c("Facility", "Total")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "300px")
```





```{r sick-calls, fig.height=40}
# Generating a linetype vector for use in the plot
plot_lines <- c(
    "solid",
    "solid",
    "dotted"
    )

# Use Color Brewer to set colors
plot_colors <- brewer.pal(3, "Paired")
plot_colors[3] <- "#000000"

# Create labels
plot_labels <- c(
    "Sick Call Requests",
    "Sick Call Encounters",
    "Total Sick Calls")

df_324_inc %>%
  
  # Subset the df to only the used cols
  select(id, facility, date,
       sick_call_requests:
       sick_call_encounters
       ) %>% 
  
  # Need the rowwise function to compute a row-at-a-time
  # in the following mutate function
  rowwise(id) %>% 
  
  # Create a new total column
  mutate(total_sick_calls = sum(c_across(
       sick_call_requests:
       sick_call_encounters
  ))) %>% 

  # Call a range of table columns and pivot long
  pivot_longer(.,
               cols=  sick_call_requests:total_sick_calls,
               names_to = "sick_call_type",
               values_to = "sick_call_count") %>%
  
  # Remove NA values
  drop_na() %>% 
  
  # Explicitly define factor levels
  mutate(sick_call_type = factor(sick_call_type, levels = c(
    "sick_call_requests",
    "sick_call_encounters",
    "total_sick_calls"
  ))) %>% 
  
  # Calling the plot and formatting
  ggplot(aes(x=date, y = sick_call_count, linetype=sick_call_type))+
  geom_line(aes(color = sick_call_type),size=.65) +
  
  # Set the linetype
  scale_linetype_manual(
    values = plot_lines,
    labels = plot_labels,
    name = "Sick Call Type:",
    guide = guide_legend(nrow = 3)
    )+
  
  # Set the color
  scale_color_manual(values = plot_colors,
    labels = plot_labels,
    name = "Sick Call Type:")+

  labs(title = "Sick Calls")+
  ylab("Number of Sick Calls")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 8),
    legend.position = "bottom"
    )+

  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =3))+
  
  # Wrap
  facet_wrap(~ facility, ncol=3)
```

## Suicide Attempts and Watches

The following shows Suicide Attempts and Suicide Watches based on the status of the data as of `r Sys.Date()`. The numbers are subject to change.

```{r}
df_suicide <- df_324_inc %>%
  
  # Select a subset of columns to work with
  select(id,
         facility,
         date,
         suicide_attempts_or_self_harm,
         suicide_watches_constant_watch_mental_health_observation) %>% 
  
  # Need the rowwise function to compute a row-at-a-time
  # in the following mutate function
  rowwise(id) %>% 
  
  # Create new total column
  mutate(total_suicide = suicide_attempts_or_self_harm +
         suicide_watches_constant_watch_mental_health_observation) %>% 
  
  # Tidy
  pivot_longer(.,
               cols= suicide_attempts_or_self_harm:
               total_suicide,
               names_to = "suicide_type",
               values_to = "suicide_count") %>% 
  
  # Remove NA
  drop_na() %>% 
  
  # Explicitly set factor levels
  mutate(segregation_type = factor(suicide_type, levels = c(
    "suicide_attempts_or_self_harm",
    "suicide_watches_constant_watch_mental_health_observation",
    "total_suicide"
  )))
```

```{r}
df_suicide %>% 
  group_by(suicide_type) %>% 
  summarise(`Total Suicide Attempt or Watch by Type` = sum(suicide_count)) %>% 
  ungroup() %>% 
  kable(caption = "Total Suicide Attempt or Watch by Type",
        col.names = c("Type", "Total")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive"))

df_suicide %>% 
  group_by(facility) %>% 
  summarise(total_suicide = sum(suicide_count)) %>% 
  arrange(desc(total_suicide)) %>% 
  ungroup() %>% 
  kable(caption = "Total Suicide Attempt or Watch by Facility",
        col.names = c("Facility", "Total")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "300px")
```


```{r suicide-facet, fig.height=40}

# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(2, "Paired")

# Create plot labels
plot_labels <- c(
      "Suicide Attempts or Self Harm",
      "Suicide Watches/Constant Watch/Mental Health Observation")

df_324_inc %>%
  
  # Subset the df to only the used cols
  select(id, facility, date,
       suicide_attempts_or_self_harm,
       suicide_watches_constant_watch_mental_health_observation
       ) %>% 
  drop_na() %>% 
 
  # Call a range of table columns and pivot long
  pivot_longer(.,
               cols=  c(suicide_attempts_or_self_harm,
       suicide_watches_constant_watch_mental_health_observation),
               names_to = "suicide_type",
               values_to = "suicide_count") %>% 
  
  # Remove NA values
  drop_na() %>% 
  
  # Explicitly define factor levels
  mutate(suicide_type = factor(suicide_type, levels = c(
    "suicide_attempts_or_self_harm",
    "suicide_watches_constant_watch_mental_health_observation"
  ))) %>% 
  
  # Calling the plot and formatting
  ggplot(aes(x=date, y = suicide_count))+
  geom_line(aes(color=suicide_type), size = .65) +
  
  # Set the color
  scale_color_manual(
    values = plot_colors,
    labels = plot_labels,
    name = "Type:"
    )+
  
  labs(title = "Suicide Attempts and Watches")+
  ylab("Number of Attempts or Watches")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 5),
    legend.position = "bottom"
    )+
  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =5))+
  
  # Wrap
  facet_wrap(~ facility, ncol = 3)


```

## Hunger Strikes

```{r hunger-strike-summary}
df_324_inc %>%
  
  # Subset the df to only the used cols
  select(id, facility, date, hunger_strikes) %>% 

  group_by(facility) %>% 
  summarise(total_hunger_strike = sum(hunger_strikes)) %>% 
  arrange(desc(total_hunger_strike)) %>% 
  ungroup() %>% 
  kable(caption = "Total Hunger Strikes by Facility",
        col.names = c("Facility", "Total Hunger Strikes")) %>% 
  kable_styling(c("hover", "striped", "condensed", "responsive")) %>% 
  scroll_box(height = "300px")
```




```{r hunger-strike-facet, fig.height=40}
# Use Color Brewer to set colors and modify
# the last color to be black for totals.
plot_colors <- brewer.pal(2, "Paired")

df_324_inc %>%
  
  # Subset the df to only the used cols
  select(id, facility, date,
       hunger_strikes
       ) %>% 
  drop_na() %>% 
 

  # Calling the plot and formatting
  ggplot(aes(x=date, y = hunger_strikes))+
  geom_line(size = .65) +
  
  # scale_color_manual(
  #   values = plot_colors,
  #   labels = c(
  #     "Suicide Attempts or Self Harm",
  #     "Suicide Watches/Constant Watch/Mental Health Observation"
  #     ),
  #   name = "Type:"
  #   )+
  # 
  labs(title = "Hunger Strikes")+
  ylab("Number of Hunger Strikes")+
  xlab("Date")+
  theme(
    strip.text = element_text(size = 5),
    legend.position = "bottom"
    )+
  
  # Set the legend to multiple rows
  guides(col = guide_legend(nrow =5))+
  
  # Wrap
  facet_wrap(~ facility, ncol = 3)
```

# Cause of Death

## Suicide as Cause of Death

Bar plot of ICE and non-ICE deaths by suicide plotted by facility as of `r Sys.Date()`.

```{r suicide-cod-plot}
df_324 %>% 
  select(id,
         facility,
         state,
         date,
         fac_operator,
         cod_suicide_ice,
         cod_suicide_not_ice) %>% 
  drop_na() %>% 
  filter(cod_suicide_ice>0|cod_suicide_not_ice >0) %>% 
  mutate(cod_suicide_total = cod_suicide_ice + cod_suicide_not_ice) %>% 
  
  # Here one could call datatable() or kable()
  # to get a list of the facilities and count
  
  # Initiate the plot
  ggplot(aes(x=cod_suicide_total, y=facility))+
  geom_boxplot()+
  scale_x_continuous(breaks = c(1,2))+
  labs(title = "Cause of Death Suicide",
       subtitle = "Includes ICE and non-ICE Deaths by Suicide")+
  xlab("COD Suicide Total")+
  ylab("Facility")
```

# Observations

Observations made on the data as they accumulate are logged below.

## 28 May 2021

1.  Adelanto ICE Processing Center - East and West have identical data for 2020, but the numbers are different for 2019 with East being much lower than west. Craig checked the inspections and the numbers are different for East and West in 2019 while they are in fact the same for 2020. Note that 2020 is a remote inspection. However, it is significant that the two are comparable in one instance and not comparable in another case.

2.  Almance County Jail and Chase County Detention Center both seem to register blank for solitary. Craig suspects this is because there were not cases of solitary reported.

## 27 May 2021

One of the important early observations has to do with the naming of facilities. ICE is often inconsistent with facility naming and the names do often change over time. Both the changing of names and inconsistent naming conventions can cause confusion as well as error when tabulating results.

1.  The graphs for "**Adelanto ICE Processing Center - East**" and "**Adelanto ICE Processing Center - West**" are identical. This is consistent with information Craig recalls seeing in one of the inspections, that the numbers for the facilities are "merged" or combined.

    -   For facet graphing consider filtering. For analysis purposes may need to treat the two facilities as "merged"

2.  "**Adelanto ICE Processing Center West**" should be combined with "**Adelanto ICE Processing Center - West**". It appears that ICE is not consistently naming the facilities, in one case the name lacks a dash character which causes them to be graphed separately.

    -   The dash was added to Adelanto West.

3.  It would be good to check if there really two facilities in Almance County. There are entries listed for "**Almance County Detention Center**" and "**Almance County Jail**".

    -   Need to verify that these are in fact distinct facilities.

4.  "**Aurora ICE Processing Center II - Annex**" and "**Aurora II Annex**" should probably be combined. Need to check the inspections to verify.

    -   "**Aurora II Annex (CO) - Jan. 29, 2021"** was changed to "**Aurora ICE Processing Center II - Annex (CO) - Jan. 29, 2021**"

# References

::: {#refs}
:::

```{r package-references, echo=FALSE}
knitr::write_bib(file = 'packages.bib')
```
